---
title: "Adding more covariates"
author: "Andy Baxter"
execute: 
  echo: false
  output: false
  warning: false
format: 
  docx:
    number-sections: true
    reference-doc: reference.docx
---

<!-- # Reading in data -->


```{r setup}
#| output: false
here::i_am("building_models/03_fuller_mcs_model_testing.qmd")
source(here::here("R", "data_import_and_cleaning.R"))

pop_data <- import_data() |> 
  data.table::as.data.table()

library(gfoRmula)
library(magrittr)
library(gFormulaMI)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)


ggplot2::theme_set(theme_sphsu_light())

# All observations, waves 8, 9 and 10

late_pop <- pop_data[wave %in% 7:10][, `:=`(
  t0 = t0 - 2,
  pre_mcs = shift(sf12mcs_dv, type = "lag"),
  pre_pcs = shift(sf12pcs_dv, type = "lag")
)][wave != 7]


intervention_pattern <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

```

# Confounders

Baseline static confounders:

 - Sex
 - Age at baseline (and age squared)
 - Ethnicity
 - Education
 - Location (`gor_dv`)

Time-varying confounders:

 - Physical health (`sf12pcs_dv`)
 - Financial distress (`econ_dist`)
 - Relationship status (`mastat_dv`)
 - Number of dependents (`dnc`)
 - Any income from benefits (`econ_benefits`)
 - Home ownership (`home_owner`)
 - Income (`log_income` or `econ_incquint`)
 
Exposures:

 - Employment (`econ_emp_bin`)



## MI method - updating MCS scores

```{r}
#| output: true
regimes <- intervention_pattern

wide_data <- as_tibble(late_pop) |> 
  select(-wave, -contains("lag")) |> 
  pivot_wider(id_cols = c(pidp, ends_with("base")), names_from = t0, values_from = c(sf12mcs_dv:dnc, pre_mcs, pre_pcs)) |> 
  select(
    sex_dv_base,
    wnw_race_base,
    hiqual_dv_base,
    age_dv_base,
    gor_dv_base,
    pre_mcs_0,
    pre_pcs_0,
    econ_emp_bin_0,
    econ_dist_0,
    dnc_0,
    econ_benefits_0,
    home_owner_0,
    log_income_0,
    pre_mcs_1,
    pre_pcs_1,
    econ_emp_bin_1,
    econ_dist_1,
    dnc_1,
    econ_benefits_1,
    home_owner_1,
    log_income_1,
    pre_mcs_2,
    pre_pcs_2,
    econ_emp_bin_2,
    econ_dist_2,
    dnc_2,
    econ_benefits_2,
    home_owner_2,
    log_income_2,
    sf12mcs_dv_2
  )

predict_matrix <- 
  matrix(
 c(
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # sex_dv_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # wnw_race_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # hiqual_dv_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # age_dv_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # gor_dv_base
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_mcs_0
1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_pcs_0
1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_emp_bin_0
1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_dist_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # dnc_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_benefits_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # home_owner_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # log_income_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_mcs_1
1, 1, 1, 1, 1, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_pcs_1
1, 1, 1, 1, 1, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_emp_bin_1
1, 1, 1, 1, 1, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_dist_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # dnc_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_benefits_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # home_owner_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # log_income_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_mcs_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_pcs_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_emp_bin_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0,  # econ_dist_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0,  # dnc_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0,  # econ_benefits_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0,  # home_owner_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0,  # log_income_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0,  # sf12mcs_dv_2
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0   # regime
), nrow = 31L, byrow = TRUE)



imputed_sets <- mice::mice(wide_data, m = 40,
                           defaultMethod = c("pmm", "logreg", "pmm", "pmm"))

imps <- gFormulaImpute(
  imputed_sets,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2"),
  trtRegimes = regimes,
  predictorMatrix = predict_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_2 ~ 0 + factor(regime))

outvals <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)

mi_results <-
  outvals |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(MI_result = Estimate,
            ll = `95% CI L`,
            ul = `95% CI U`,
            intervention = regimes_t$intervention)
```

```{r}
#| output: true
  ggplot(mi_results, aes(MI_result, intervention, xmin = ll, xmax = ul)) +
    geom_point(size = 2)  +
    geom_errorbar(width = 0.3) +
  scale_colour_sphsu()
```



## MI method - static base MCS score

```{r}
#| output: false
regimes <- intervention_pattern

wide_data <- as_tibble(late_pop) |> 
  select(-wave, -contains("lag")) |> 
  pivot_wider(id_cols = c(pidp, ends_with("base")), names_from = t0, values_from = c(sf12mcs_dv:dnc, pre_mcs, pre_pcs)) |> 
  select(
    sex_dv_base,
    wnw_race_base,
    hiqual_dv_base,
    age_dv_base,
    gor_dv_base,
    sf12mcs_dv_base,
    pre_pcs_0,
    econ_emp_bin_0,
    econ_dist_0,
    dnc_0,
    econ_benefits_0,
    home_owner_0,
    log_income_0,
    pre_pcs_1,
    econ_emp_bin_1,
    econ_dist_1,
    dnc_1,
    econ_benefits_1,
    home_owner_1,
    log_income_1,
    pre_pcs_2,
    econ_emp_bin_2,
    econ_dist_2,
    dnc_2,
    econ_benefits_2,
    home_owner_2,
    log_income_2,
    sf12mcs_dv_2
  )

predict_matrix <- 
  matrix(
 c(
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # sex_dv_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # wnw_race_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # hiqual_dv_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # age_dv_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # gor_dv_base
0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # sf12_mcs_base
1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_pcs_0
1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_emp_bin_0
1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_dist_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # dnc_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_benefits_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # home_owner_0
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # log_income_0
1, 1, 1, 1, 1, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_pcs_1
1, 1, 1, 1, 1, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_emp_bin_1
1, 1, 1, 1, 1, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_dist_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # dnc_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_benefits_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # home_owner_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # log_income_1
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0,  # pre_pcs_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0,  # econ_emp_bin_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0,  # econ_dist_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0,  # dnc_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0,  # econ_benefits_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0,  # home_owner_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0,  # log_income_2
1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0,  # sf12mcs_dv_2
1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0  # regime
), nrow = 29L, byrow = TRUE)



imputed_sets <- mice::mice(wide_data, m = 40,
                           defaultMethod = c("pmm", "logreg", "pmm", "pmm"))

imps <- gFormulaImpute(
  imputed_sets,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2"),
  trtRegimes = regimes,
  predictorMatrix = predict_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_2 ~ 0 + factor(regime))

outvals <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)

mi_results_2 <-
  outvals |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(MI_result = Estimate,
            ll = `95% CI L`,
            ul = `95% CI U`,
            intervention = regimes_t$intervention)
```

```{r}
#| output: true
  ggplot(mi_results_2, aes(MI_result, intervention, xmin = ll, xmax = ul)) +
    geom_point(size = 2)  +
    geom_errorbar(width = 0.3) +
  scale_colour_sphsu()
```


