---
title: "All time periods"
format: html
---

```{r setup}
#| output: false
here::i_am("building_models/06_all_time_points.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

pop_data <- import_data() |> 
  data.table::as.data.table()

library(gfoRmula)
library(magrittr)
library(gFormulaMI)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)


ggplot2::theme_set(theme_sphsu_light())


late_pop <- pop_data[, `:=`(
  t0 = t0 - 1,
  pre_mcs = shift(sf12mcs_dv, type = "lag"),
  pre_pcs = shift(sf12pcs_dv, type = "lag")
)][wave != 1]

intervention_pattern <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

```

# Confounders

Baseline static confounders:

 - Sex
 - Age at baseline (and age squared)
 - Ethnicity
 - Education
 - Location (`gor_dv`)

Time-varying confounders:

 - Physical health (`sf12pcs_dv`)
 - Financial distress (`econ_dist`)
 - Relationship status (`mastat_dv`)
 - Number of dependents (`dnc`)
 - Any income from benefits (`econ_benefits`)
 - Home ownership (`home_owner`)
 - Income (`log_income` or `econ_incquint`)
 
Exposures:

 - Employment at waves 7, 8, 9 and 10 (`econ_emp_bin`)

# Method 1: updating MCS
 
```{r}
#| output: false
regimes <- intervention_pattern

wide_data <- late_pop |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  wnw_race_base,
                  hiqual_dv_base,
                  age_dv_base,
                  gor_dv_base),
    pre_mcs,
    pre_pcs,
    econ_dist,
    dnc,
    econ_benefits,
    home_owner,
    econ_emp_bin,
    log_income
  )


predictor_matrix <- make_predictor_matrix(wide_data)

imputed_sets <- mice::mice(wide_data, m = 20,
                           defaultMethod = c("pmm", "logreg", "pmm", "pmm"))

imps <- gFormulaImpute(
  imputed_sets,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2"),
  trtRegimes = regimes,
  predictorMatrix = predict_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_2 ~ 0 + factor(regime))

outvals <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

