---
title: "Full models"
author: "Andy Baxter"
execute: 
  output: false
---

```{r setup}
#| output: false
here::i_am("building_models/07_full_models.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(magrittr)
library(gFormulaMI)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)

pop_data <- import_data() |> 
  select(-econ_incquint) 


ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 2 (baseline), and wave 6

pop_dt <- pop_data |> 
  filter(wave == 2, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(wave == 6, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(!is.na(mastat_dv_base)) |> 
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |> 
  data.table::as.data.table()

late_pop <- pop_dt[, `:=`(
  t0 = t0 - 2,
  pre_mcs = shift(sf12mcs_dv, type = "lag"),
  pre_pcs = shift(sf12pcs_dv, type = "lag")
)][wave %in% 3:6 & response == 1]

no_mh_outcomes <- late_pop |> 
  group_by(pidp) |> 
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |> 
  filter(n_missing == 4, base_missing) |> 
  select(pidp) |> 
  as.data.table()

late_pop <- late_pop[!no_mh_outcomes, on = .(pidp)] 

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

nsamples <- 1000

imputed_late <- mice::mice(late_pop[,-c(1:3)], defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 20)

```

# Confounders

Baseline static confounders:

 - Sex
 - Age at baseline (and age squared)
 - Ethnicity
 - Education
 - Location (`gor_dv`)

Time-varying confounders:

 - Physical health (`sf12pcs_dv`)
 - Financial distress (`econ_dist`)
 - Relationship status (`mastat_dv`)
 - Number of dependants (`dnc`)
 - Any income from benefits (`econ_benefits`)
 - Home ownership (`home_owner`)
 - Income (`log_income`)
 
Exposures:

 - Employment at waves 3, 4, 5, 6 (`econ_emp_bin`)

# Four periods of intervention

## Updating MCS as time-varying confounder

### gformula

```{r 4dynamic-gform}
outcome_type <- "continuous_eof"
outcome_name <- "sf12mcs_dv"
intvars <- c("econ_emp_bin")

time_points <- 4


covnames <-
  c(
    "pre_mcs",
    "pre_pcs",
    "econ_emp_bin",
    "econ_dist",
    "dnc",
    "econ_benefits",
    "home_owner",
    "log_income"
  )


covtypes <- c(
  "normal", # pre_mcs
  "normal", # pre_pcs
  "binary", # econ_emp_bin
  "binary", # econ_dist
  "categorical", # dnc
  "binary", # econ_benefits
  "binary", # home_owner
  "normal" # log_income
)

basecovs <- c(
  "sex_dv_base",
  "home_owner_base",
  "wnw_race_base",
  "hiqual_dv_base",
  "age_dv_base"
  # "dnc_base",
  # "mastat_dv_base",
)

ymodel <- sf12mcs_dv ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +  # Baseline - sex
  pre_mcs + pre_pcs + econ_dist + dnc + econ_benefits + home_owner +     # Time-varying confounders
  econ_emp_bin + log_income    # Time-varying exposures


covparams <- list(
  covmodels = c(
    pre_mcs ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_pre_mcs + lag1_pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + 
      lag1_econ_emp_bin + lag1_log_income,
    pre_pcs ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_pre_pcs + pre_mcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      lag1_econ_emp_bin + lag1_log_income,
    econ_emp_bin ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + lag1_log_income + lag1_econ_emp_bin,
    econ_dist ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    dnc ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + econ_emp_bin + lag1_log_income,
    econ_benefits ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    home_owner ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + econ_dist + dnc + econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    log_income ~
      sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_log_income + pre_pcs + pre_mcs + econ_dist + dnc + econ_benefits + home_owner + econ_emp_bin
  )
)






lagged_variables <- c(
  "pre_pcs",
  "pre_mcs",
  "econ_dist",
  "dnc",
  "home_owner",
  "econ_benefits",
  "econ_emp_bin",
  "log_income"
)


histories <- c(lagged)
histvars <- list(lagged_variables)

ncores <- min(32, parallel::detectCores() - 1)



interventions <- intervention_pattern_4 |>
  map( ~ list(c(static, c(.x))))

intvars <- as.list(rep("econ_emp_bin", length(interventions)))

int_description <-
  map(interventions, ~paste(.x[[1]][-1], collapse = "-") |> paste0("econ_emp_bin:", x = _))


gform_cont_eof <- gformula(obs_data = late_pop,
                           outcome_type = outcome_type,
                           id = "pidp",
                           time_name = "t0",
                           time_points = time_points,
                           covnames = covnames,
                           outcome_name = outcome_name,
                           covtypes = covtypes,
                           covparams = covparams,
                           # covpredict_custom = covpredict_custom,
                           ymodel = ymodel,
                           
                           intvars = intvars,
                           interventions = interventions,
                           int_descript = int_description,
                           
                           # covfits_custom = c(NA, NA, NA, NA, incr_fit),
                           # covpredict_custom = covpredict_custom,
                           
                           histories = histories,
                           histvars = histvars,
                           basecovs = basecovs,
                           seed = 1234,
                           parallel = TRUE,
                           nsamples = nsamples, # Number of bootstrap samples
                           ncores = ncores,
                           ref_int = 1
)


```
 
### gFormulaMI

```{r 4dynamic-mi}
regimes <- intervention_pattern_4

wide_data <- late_pop |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  wnw_race_base,
                  hiqual_dv_base,
                  age_dv_base,
                  gor_dv_base),
    outcome = sf12mcs_dv,
    pre_mcs,
    pre_pcs,
    econ_dist,
    dnc,
    econ_benefits,
    home_owner,
    econ_emp_bin,
    log_income
  )


predictor_matrix <- make_predictor_matrix(wide_data)

imputed_sets <- mice::mice(wide_data, m = 20,
                           defaultMethod = c("pmm", "logreg", "pmm", "pmm"))

imps <- gFormulaImpute(
  imputed_sets,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

outvals <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

### Comparing results

```{r 4dynamic-tab}
#| output: true
mod_4dynamic <- outvals |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(MI_result = Estimate,
            MI_ll = `95% CI L`,
            MI_ul = `95% CI U`) |>
  bind_cols(
    gform_cont_eof$result |>
      filter(`Interv.` != 0) |>
      select(
        gform_result = `g-form mean`,
        gform_ll = `Mean lower 95% CI`,
        gform_ul = `Mean upper 95% CI`
      )
  ) |>
  bind_cols(regimes_t)

mod_4dynamic |> 
  knitr::kable(digits = 3)
  
```

```{r 4dynamic-graph}
#| output: true
mod_4dynamic |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(result, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```

## MCS baseline as static confounder

### gformula

```{r 4baseline-gform}
outcome_type <- "continuous_eof"
outcome_name <- "sf12mcs_dv"
intvars <- c("econ_emp_bin")

time_points <- 4


covnames <-
  c(
    "pre_pcs",
    "econ_emp_bin",
    "econ_dist",
    "dnc",
    "econ_benefits",
    "home_owner",
    "log_income"
  )


covtypes <- c(
  "normal", # pre_pcs
  "binary", # econ_emp_bin
  "binary", # econ_dist
  "categorical", # dnc
  "binary", # econ_benefits
  "binary", # home_owner
  "normal" # log_income
)

basecovs <- c(
  "sf12mcs_dv_base",
  "sex_dv_base",
  "home_owner_base",
  "wnw_race_base",
  "hiqual_dv_base",
  "age_dv_base"
  # "dnc_base",
  # "mastat_dv_base",
)

ymodel <- sf12mcs_dv ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +  
  pre_pcs + econ_dist + dnc + econ_benefits + home_owner +     # Time-varying confounders
  econ_emp_bin + log_income    # Time-varying exposures


covparams <- list(
  covmodels = c(
    pre_pcs ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      lag1_pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      lag1_econ_emp_bin + lag1_log_income,
    econ_emp_bin ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + lag1_log_income + lag1_econ_emp_bin,
    econ_dist ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    dnc ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + econ_emp_bin + lag1_log_income,
    econ_benefits ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    home_owner ~ sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + econ_dist + dnc + econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    log_income ~
      sex_dv_base + home_owner_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      lag1_log_income + pre_pcs + econ_dist + dnc + econ_benefits + home_owner + econ_emp_bin
  )
)






lagged_variables <- c(
  "pre_pcs",
  "econ_dist",
  "dnc",
  "home_owner",
  "econ_benefits",
  "econ_emp_bin",
  "log_income"
)


histories <- c(lagged)
histvars <- list(lagged_variables)


interventions <- intervention_pattern_4 |>
  map( ~ list(c(static, c(.x))))

intvars <- as.list(rep("econ_emp_bin", length(interventions)))

int_description <-
  map(interventions, ~paste(.x[[1]][-1], collapse = "-") |> paste0("econ_emp_bin:", x = _))


gform_cont_eof <- gformula(obs_data = late_pop,
                           outcome_type = outcome_type,
                           id = "pidp",
                           time_name = "t0",
                           time_points = time_points,
                           covnames = covnames,
                           outcome_name = outcome_name,
                           covtypes = covtypes,
                           covparams = covparams,
                           # covpredict_custom = covpredict_custom,
                           ymodel = ymodel,
                           
                           intvars = intvars,
                           interventions = interventions,
                           int_descript = int_description,
                           
                           # covfits_custom = c(NA, NA, NA, NA, incr_fit),
                           # covpredict_custom = covpredict_custom,
                           
                           histories = histories,
                           histvars = histvars,
                           basecovs = basecovs,
                           seed = 1234,
                           parallel = TRUE,
                           nsamples = nsamples, # Number of bootstrap samples
                           ncores = ncores,
                           ref_int = 1
)
```
 
### gFormulaMI

```{r 4baseline-mi}
regimes <- intervention_pattern_4

wide_data <- late_pop |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  wnw_race_base,
                  hiqual_dv_base,
                  age_dv_base,
                  gor_dv_base,
                  sf12mcs_dv_base),
    outcome = sf12mcs_dv,
    pre_pcs,
    econ_dist,
    dnc,
    econ_benefits,
    home_owner,
    econ_emp_bin,
    log_income
  )


predictor_matrix <- make_predictor_matrix(wide_data)

imputed_sets <- mice::mice(wide_data, m = 20,
                           defaultMethod = c("pmm", "logreg", "pmm", "pmm"))

imps <- gFormulaImpute(
  imputed_sets,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

outvals <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

### Comparing results

```{r 4baseline-tab}
#| output: true
mod_4baseline <- outvals |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(MI_result = Estimate,
            MI_ll = `95% CI L`,
            MI_ul = `95% CI U`) |>
  bind_cols(
    gform_cont_eof$result |>
      filter(`Interv.` != 0) |>
      select(
        gform_result = `g-form mean`,
        gform_ll = `Mean lower 95% CI`,
        gform_ul = `Mean upper 95% CI`
      )
  ) |>
  bind_cols(regimes_t)

mod_4baseline |> 
  knitr::kable(digits = 3)
  
```

```{r 4baseline-graph}
#| output: true
mod_4baseline |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(result, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```