---
title: "Full models"
author: "Andy Baxter"
format: html
execute: 
  output: false
---

```{r setup}
#| output: false
here::i_am("building_models/07_full_models.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(magrittr)
library(gFormulaMI)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(mice)

pop_data <- import_data() |> 
  select(-econ_incquint) 


ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 2 (baseline), and wave 6

pop_dt <- pop_data |> 
  filter(wave == 2, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(wave == 6, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(!is.na(mastat_dv_base)) |> 
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |> 
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 2,
  pre_mcs = shift(sf12mcs_dv, type = "lag"),
  pre_pcs = shift(sf12pcs_dv, type = "lag")
)][wave %in% 3:6]

no_mh_outcomes <- long_data |> 
  group_by(pidp) |> 
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |> 
  filter(n_missing == 4, base_missing) |> 
  select(pidp) |> 
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)] 

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

nsamples <- 500
```

# New plan

- make all useable variables in wide format
  - with econ_emp_bin as factor
- impute 20 pops across all variables
- complete and pivot for gformula calculations
- pass long objec to gformula calculations
  - with econ_emp-bin as binary
- pass mids object to MI method, but amend predictor matrix

Combine via Rubin's rules:

Combined mean:

$$
\bar{\theta} = \frac{1}{m}\left (\sum_{i=1}^m{\theta_i}\right )
$$

Within-imputation pooled variance:

$$
V_W = \frac{1}{m} \sum_{i=1}^m{SE_i^2}
$$

Between-imputation variance:

$$
V_B = \frac{\sum_{i=1}^m (\theta_i - \overline{\theta})^2}{m-1}
$$

Combined variance:

$$
V_{Total} = V_W + V_B + \frac{V_B}{m}
$$


```{r impute-two-datasets}
wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |> 
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  wnw_race_base,
                  hiqual_dv_base,
                  age_dv_base,
                  gor_dv_base,
                  sf12mcs_dv_base),
    outcome = sf12mcs_dv,
    pre_mcs,
    pre_pcs,
    econ_dist,
    dnc,
    econ_benefits,
    home_owner,
    econ_emp_bin,
    log_income
  )

wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 30, maxit = 10)

imputed_wide <- mice::complete(wide_mids, action = "long", include = TRUE)

imputed_long <- imputed_wide |>
  pivot_longer(
    matches("*_\\d$"),
    names_to = c(".value", "t0"),
    names_pattern = "(.*)_(\\d)",
    names_transform = list(t0 = as.integer)
  ) |>
  # (\(x) cbind(long_data[,1:3], x))()
  nest(data = -.imp) |> 
  mutate(data = map(data, as.data.table))

base_predictor_matrix <- make_predictor_matrix(wide_data)

```


```{r old-method-impute-long}
#| include: false
# imputed_late <- long_data[,-c(1:3)] |> 
#   mutate(econ_emp_bin = as_factor(econ_emp_bin)) |> 
#   mice::mice(defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 20)
# 
# full_impute <- mice::complete(imputed_late, action = "long", include = TRUE)


```


# Confounders

Baseline static confounders:

 - Sex
 - Age at baseline (and age squared)
 - Ethnicity
 - Education
 - Location (`gor_dv`)

Time-varying confounders:

 - Mental health (in updating models; `sf12mcs_dv`)
 - Physical health (`sf12pcs_dv`)
 - Financial distress (`econ_dist`)
 - Relationship status (`mastat_dv`)
 - Number of dependants (`dnc`)
 - Any income from benefits (`econ_benefits`)
 - Home ownership (`home_owner`)
 - Income (`log_income`)
 
Exposures:

 - Employment at waves 3, 4, 5, 6 (`econ_emp_bin`)

# Four periods of intervention

## Updating MCS as time-varying confounder

### gformula

```{r 4dynamic-gform}
outcome_type <- "continuous_eof"
outcome_name <- "sf12mcs_dv"
intvars <- c("econ_emp_bin")

time_points <- 4


covnames <-
  c(
    "pre_mcs",
    "pre_pcs",
    "econ_emp_bin",
    "econ_dist",
    "dnc",
    "econ_benefits",
    "home_owner",
    "log_income"
  )


covtypes <- c(
  "normal", # pre_mcs
  "normal", # pre_pcs
  "binary", # econ_emp_bin
  "categorical", # econ_dist
  "categorical", # dnc
  "categorical", # econ_benefits
  "categorical", # home_owner
  "normal" # log_income
)

basecovs <- c(
  "sex_dv_base",
  "gor_dv_base",
  "wnw_race_base",
  "hiqual_dv_base",
  "age_dv_base"
  # "dnc_base",
  # "mastat_dv_base",
)

ymodel <- sf12mcs_dv ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +  # Baseline - sex
  pre_mcs + pre_pcs + econ_dist + dnc + econ_benefits + home_owner +     # Time-varying confounders
  econ_emp_bin + log_income    # Time-varying exposures


covparams <- list(
  covmodels = c(
    pre_mcs ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_pre_mcs + lag1_pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + 
      lag1_econ_emp_bin + lag1_log_income,
    pre_pcs ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_pre_pcs + pre_mcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      lag1_econ_emp_bin + lag1_log_income,
    econ_emp_bin ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + lag1_log_income + lag1_econ_emp_bin,
    econ_dist ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    dnc ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + econ_emp_bin + lag1_log_income,
    econ_benefits ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    home_owner ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + econ_dist + dnc + econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    log_income ~
      sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_log_income + pre_pcs + pre_mcs + econ_dist + dnc + econ_benefits + home_owner + econ_emp_bin
  )
)






lagged_variables <- c(
  "pre_pcs",
  "pre_mcs",
  "econ_dist",
  "dnc",
  "home_owner",
  "econ_benefits",
  "econ_emp_bin",
  "log_income"
)


histories <- c(lagged)
histvars <- list(lagged_variables)

ncores <- min(32, parallel::detectCores() - 1)



interventions <- intervention_pattern_4 |>
  map( ~ list(c(static, c(.x))))

intvars <- as.list(rep("econ_emp_bin", length(interventions)))

int_description <-
  map(interventions, ~paste(.x[[1]][-1], collapse = "-") |> paste0("econ_emp_bin:", x = _))

gform_long <- imputed_long |>
  mutate(gform_out = map(data, function(data) {
    
    data$econ_emp_bin <- as.integer(as.character(data$econ_emp_bin))
    
    gform_cont_eof <- gformula(
      obs_data = data,
      outcome_type = outcome_type,
      id = ".id",
      time_name = "t0",
      time_points = time_points,
      covnames = covnames,
      outcome_name = outcome_name,
      covtypes = covtypes,
      covparams = covparams,
      ymodel = ymodel,
      
      intvars = intvars,
      interventions = interventions,
      int_descript = int_description,
      
      
      histories = histories,
      histvars = histvars,
      basecovs = basecovs,
      seed = 1234,
      parallel = TRUE,
      nsamples = nsamples,
      ncores = ncores,
      ref_int = 1
    )
    
    gform_cont_eof$result
    
  }))


gform_4dynamic <- pool_gform_estimates(gform_long)
```
 
### gFormulaMI

```{r 4dynamic-mi}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

predictor_matrix["sf12mcs_dv_base",] <- 0
predictor_matrix[,"sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4dynamic <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

### Comparing results

```{r 4dynamic-tab}
#| output: true
mod_4dynamic <- mi_4dynamic |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(
    gform_4dynamic |> filter(interv != 0) |> select()
  ) |>
  bind_cols(regimes_t)

mod_4dynamic |> 
  knitr::kable(digits = 3)
  
```

```{r 4dynamic-graph}
#| output: true
mod_4dynamic |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```

## MCS baseline as static confounder

### gformula

```{r 4baseline-gform}
outcome_type <- "continuous_eof"
outcome_name <- "sf12mcs_dv"
intvars <- c("econ_emp_bin")

time_points <- 4


covnames <-
  c(
    "pre_pcs",
    "econ_emp_bin",
    "econ_dist",
    "dnc",
    "econ_benefits",
    "home_owner",
    "log_income"
  )


covtypes <- c(
  "normal", # pre_pcs
  "binary", # econ_emp_bin
  "categorical", # econ_dist
  "categorical", # dnc
  "categorical", # econ_benefits
  "categorical", # home_owner
  "normal" # log_income
)

basecovs <- c(
  "sf12mcs_dv_base",
  "sex_dv_base",
  "gor_dv_base",
  "wnw_race_base",
  "hiqual_dv_base",
  "age_dv_base"
  # "dnc_base",
  # "mastat_dv_base",
)

ymodel <- sf12mcs_dv ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +  
  pre_pcs + econ_dist + dnc + econ_benefits + home_owner +     # Time-varying confounders
  econ_emp_bin + log_income    # Time-varying exposures


covparams <- list(
  covmodels = c(
    pre_pcs ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      lag1_pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      lag1_econ_emp_bin + lag1_log_income,
    econ_emp_bin ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + lag1_log_income + lag1_econ_emp_bin,
    econ_dist ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    dnc ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + econ_emp_bin + lag1_log_income,
    econ_benefits ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    home_owner ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + econ_dist + dnc + econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    log_income ~
      sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      lag1_log_income + pre_pcs + econ_dist + dnc + econ_benefits + home_owner + econ_emp_bin
  )
)






lagged_variables <- c(
  "pre_pcs",
  "econ_dist",
  "dnc",
  "home_owner",
  "econ_benefits",
  "econ_emp_bin",
  "log_income"
)


histories <- c(lagged)
histvars <- list(lagged_variables)


interventions <- intervention_pattern_4 |>
  map( ~ list(c(static, c(.x))))

intvars <- as.list(rep("econ_emp_bin", length(interventions)))

int_description <-
  map(interventions, ~paste(.x[[1]][-1], collapse = "-") |> paste0("econ_emp_bin:", x = _))


gform_long <- imputed_long |>
  mutate(gform_out = map(data, function(data) {
    
    data$econ_emp_bin <- as.integer(as.character(data$econ_emp_bin))
    
    gform_cont_eof <- gformula(
      obs_data = data,
      outcome_type = outcome_type,
      id = ".id",
      time_name = "t0",
      time_points = time_points,
      covnames = covnames,
      outcome_name = outcome_name,
      covtypes = covtypes,
      covparams = covparams,
      ymodel = ymodel,
      
      intvars = intvars,
      interventions = interventions,
      int_descript = int_description,
      
      
      histories = histories,
      histvars = histvars,
      basecovs = basecovs,
      seed = 1234,
      parallel = TRUE,
      nsamples = nsamples,
      ncores = ncores,
      ref_int = 1
    )
    
    gform_cont_eof$result
    
  }))

gform_4baseline <- pool_gform_estimates(gform_long)
```
 
### gFormulaMI

```{r 4baseline-mi}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

predictor_matrix[which(grepl("pre_mcs_\\d", rownames(predictor_matrix))),] <- 0
predictor_matrix[,which(grepl("pre_mcs_\\d", colnames(predictor_matrix)))] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4baseline <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

### Comparing results

```{r 4baseline-tab}
#| output: true
mod_4baseline <- mi_4baseline |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(
    gform_4baseline |> filter(interv != 0) |> select()
  ) |>
  bind_cols(regimes_t)

mod_4baseline |> 
  knitr::kable(digits = 3)
  
```

```{r 4baseline-graph}
#| output: true
mod_4baseline |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```

# Three periods of intervention

## Updating MCS as time-varying confounder

### gformula

```{r 3dynamic-gform}
outcome_type <- "continuous_eof"
outcome_name <- "sf12mcs_dv"
intvars <- c("econ_emp_bin")

time_points <- 4


covnames <-
  c(
    "pre_mcs",
    "pre_pcs",
    "econ_emp_bin",
    "econ_dist",
    "dnc",
    "econ_benefits",
    "home_owner",
    "log_income"
  )


covtypes <- c(
  "normal", # pre_mcs
  "normal", # pre_pcs
  "binary", # econ_emp_bin
  "categorical", # econ_dist
  "categorical", # dnc
  "categorical", # econ_benefits
  "categorical", # home_owner
  "normal" # log_income
)

basecovs <- c(
  "sex_dv_base",
  "gor_dv_base",
  "wnw_race_base",
  "hiqual_dv_base",
  "age_dv_base"
  # "dnc_base",
  # "mastat_dv_base",
)

ymodel <- sf12mcs_dv ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +  # Baseline - sex
  pre_mcs + pre_pcs + econ_dist + dnc + econ_benefits + home_owner +     # Time-varying confounders
  econ_emp_bin + log_income    # Time-varying exposures


covparams <- list(
  covmodels = c(
    pre_mcs ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_pre_mcs + lag1_pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + 
      lag1_econ_emp_bin + lag1_log_income,
    pre_pcs ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_pre_pcs + pre_mcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      lag1_econ_emp_bin + lag1_log_income,
    econ_emp_bin ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + lag1_log_income + lag1_econ_emp_bin,
    econ_dist ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    dnc ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + econ_emp_bin + lag1_log_income,
    econ_benefits ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + lag1_econ_dist + dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    home_owner ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      pre_mcs + pre_pcs + econ_dist + dnc + econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    log_income ~
      sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base +
      lag1_log_income + pre_pcs + pre_mcs + econ_dist + dnc + econ_benefits + home_owner + econ_emp_bin
  )
)






lagged_variables <- c(
  "pre_pcs",
  "pre_mcs",
  "econ_dist",
  "dnc",
  "home_owner",
  "econ_benefits",
  "econ_emp_bin",
  "log_income"
)


histories <- c(lagged)
histvars <- list(lagged_variables)

ncores <- min(32, parallel::detectCores() - 1)



interventions <- intervention_pattern_3 |>
  map(~ list(c(set_n_periods, NA_integer_, .x)))

intvars <- as.list(rep("econ_emp_bin", length(interventions)))

int_description <-
  map(
    interventions,
    ~ paste(.x[[1]][-1], collapse = "-") |>
      str_remove_all("NA-") |>
      paste0("econ_emp_bin:", x = _)
  )

gform_long <- imputed_long |>
  mutate(gform_out = map(data, function(data) {
    
    data$econ_emp_bin <- as.integer(as.character(data$econ_emp_bin))
    
    gform_cont_eof <- gformula(
      obs_data = data,
      outcome_type = outcome_type,
      id = ".id",
      time_name = "t0",
      time_points = time_points,
      covnames = covnames,
      outcome_name = outcome_name,
      covtypes = covtypes,
      covparams = covparams,
      ymodel = ymodel,
      
      intvars = intvars,
      interventions = interventions,
      int_descript = int_description,
      
      
      histories = histories,
      histvars = histvars,
      basecovs = basecovs,
      seed = 1234,
      parallel = TRUE,
      nsamples = nsamples,
      ncores = ncores,
      ref_int = 1
    )
    
    gform_cont_eof$result
    
  }))


gform_3dynamic <- pool_gform_estimates(gform_long)
```
 
### gFormulaMI

```{r 3dynamic-mi}
regimes <- intervention_pattern_3

predictor_matrix <- base_predictor_matrix

predictor_matrix["sf12mcs_dv_base",] <- 0
predictor_matrix[,"sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_3dynamic <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

### Comparing results

```{r 3dynamic-tab}
#| output: true
mod_3dynamic <- mi_3dynamic |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(
    gform_3dynamic |> filter(interv != 0) |> select()
  ) |>
  bind_cols(regimes_t)

mod_3dynamic |> 
  knitr::kable(digits = 3)
  
```

```{r 3dynamic-graph}
#| output: true
mod_3dynamic |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```

## MCS baseline as static confounder

### gformula

```{r 3baseline-gform}
outcome_type <- "continuous_eof"
outcome_name <- "sf12mcs_dv"
intvars <- c("econ_emp_bin")

time_points <- 4


covnames <-
  c(
    "pre_pcs",
    "econ_emp_bin",
    "econ_dist",
    "dnc",
    "econ_benefits",
    "home_owner",
    "log_income"
  )


covtypes <- c(
  "normal", # pre_pcs
  "binary", # econ_emp_bin
  "categorical", # econ_dist
  "categorical", # dnc
  "categorical", # econ_benefits
  "categorical", # home_owner
  "normal" # log_income
)

basecovs <- c(
  "sf12mcs_dv_base",
  "sex_dv_base",
  "gor_dv_base",
  "wnw_race_base",
  "hiqual_dv_base",
  "age_dv_base"
  # "dnc_base",
  # "mastat_dv_base",
)

ymodel <- sf12mcs_dv ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +  
  pre_pcs + econ_dist + dnc + econ_benefits + home_owner +     # Time-varying confounders
  econ_emp_bin + log_income    # Time-varying exposures


covparams <- list(
  covmodels = c(
    pre_pcs ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      lag1_pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      lag1_econ_emp_bin + lag1_log_income,
    econ_emp_bin ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + lag1_log_income + lag1_econ_emp_bin,
    econ_dist ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    dnc ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + econ_dist + lag1_dnc + lag1_econ_benefits + lag1_home_owner + econ_emp_bin + lag1_log_income,
    econ_benefits ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + lag1_econ_dist + dnc + lag1_econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    home_owner ~ sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      pre_pcs + econ_dist + dnc + econ_benefits + lag1_home_owner +
      econ_emp_bin + lag1_log_income,
    log_income ~
      sex_dv_base + gor_dv_base + wnw_race_base + hiqual_dv_base + age_dv_base + sf12mcs_dv_base +
      lag1_log_income + pre_pcs + econ_dist + dnc + econ_benefits + home_owner + econ_emp_bin
  )
)






lagged_variables <- c(
  "pre_pcs",
  "econ_dist",
  "dnc",
  "home_owner",
  "econ_benefits",
  "econ_emp_bin",
  "log_income"
)


histories <- c(lagged)
histvars <- list(lagged_variables)


interventions <- intervention_pattern_3 |>
  map(~ list(c(set_n_periods, NA_integer_, .x)))

intvars <- as.list(rep("econ_emp_bin", length(interventions)))

int_description <-
  map(
    interventions,
    ~ paste(.x[[1]][-1], collapse = "-") |>
      str_remove_all("NA-") |>
      paste0("econ_emp_bin:", x = _)
  )


gform_long <- imputed_long |>
  mutate(gform_out = map(data, function(data) {
    
    data$econ_emp_bin <- as.integer(as.character(data$econ_emp_bin))
    
    gform_cont_eof <- gformula(
      obs_data = data,
      outcome_type = outcome_type,
      id = ".id",
      time_name = "t0",
      time_points = time_points,
      covnames = covnames,
      outcome_name = outcome_name,
      covtypes = covtypes,
      covparams = covparams,
      ymodel = ymodel,
      
      intvars = intvars,
      interventions = interventions,
      int_descript = int_description,
      
      
      histories = histories,
      histvars = histvars,
      basecovs = basecovs,
      seed = 1234,
      parallel = TRUE,
      nsamples = nsamples,
      ncores = ncores,
      ref_int = 1
    )
    
    gform_cont_eof$result
    
  }))

gform_3baseline <- pool_gform_estimates(gform_long)
```
 
### gFormulaMI

```{r 3baseline-mi}
regimes <- intervention_pattern_3

predictor_matrix <- base_predictor_matrix

predictor_matrix[which(grepl("pre_mcs_\\d", rownames(predictor_matrix))),] <- 0
predictor_matrix[,which(grepl("pre_mcs_\\d", colnames(predictor_matrix)))] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_3baseline <- syntheticPool(fits)

regimes_t <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

### Comparing results

```{r 3baseline-tab}
#| output: true
mod_3baseline <- mi_3baseline |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(
    gform_3baseline |> filter(interv != 0) |> select()
  ) |>
  bind_cols(regimes_t)

mod_3baseline |> 
  knitr::kable(digits = 3)
  
```

```{r 3baseline-graph}
#| output: true
mod_3baseline |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```