---
title: "Estimating three-period with baseline and accumulating variables"
author: "Darwin Del Castillo"
format: 
  html:
    toc: true
    fig-format: svg
execute: 
  output: false
---

# Setup

```{r setup}
#| output: false
here::i_am("building_models/10_three_waves.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(gFormulaMI)
library(magrittr)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(DiagrammeR)

pop_data <- import_data() |> 
  select(-econ_incquint) 

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 2 (baseline), and wave 6

pop_dt <- pop_data |> 
  filter(wave == 2, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(wave == 6, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(!is.na(mastat_dv_base)) |> 
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |> 
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 2,
  pcs_lagged = shift(sf12pcs_dv, type = "lag"), # when it acts as a confounder, it should be lagged, when it acts as an outcome, it should not be lagged
# mcs_lagged = shift(sf12mcs_dv, type = "lag"), # when it acts as a confounder, it should be lagged, when it acts as an outcome, it should not be lagged
# the other time-varying confounders should also be lagged
  lagged_dnc = shift(dnc, type = "lag"),
  lagged_home_owner = shift(home_owner, type = "lag"),
  lagged_econ_benefits = shift(econ_benefits, type = "lag"),
  mcs_outcome = sf12mcs_dv
)][wave %in% 3:6]

# filtering out those with no observed MCS outcomes (at least 2 missing, and baseline missing)
no_mh_outcomes <- long_data |> 
  group_by(pidp) |> 
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |> 
  filter(n_missing >= 2, base_missing) |> 
  select(pidp) |> 
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)] 

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()
```

# Three periods of intervention - with intermediate outcomes

```{r model-cummulative3-dag}
#| echo: false
#| output: true
grViz("digraph {
  graph []
  node [fontname=Arial]
  P0 [label=<PCS<SUB>1</SUB>>]
  P1 [label=<PCS<SUB>2</SUB>>]
  P2 [label=<PCS<SUB>3</SUB>>]
  E0 [label=<Empl<SUB>1</SUB>>]
  E1 [label=<Empl<SUB>2</SUB>>]
  E2 [label=<Empl<SUB>3</SUB>>]
  I0 [label=<Inc<SUB>1</SUB>>]
  I1 [label=<Inc<SUB>2</SUB>>]
  I2 [label=<Inc<SUB>3</SUB>>]
  C0 [label=<Conf<SUP>Int</SUP><SUB>1</SUB>>]
  C1 [label=<Conf<SUP>Int</SUP><SUB>2</SUB>>]
  C2 [label=<Conf<SUP>Int</SUP><SUB>3</SUB>>]
  Y0 [label=<<I>MCS<SUB>1</SUB></I>>]
  Y1 [label=<<I>MCS<SUB>2</SUB></I>>]
  Y2 [label=<<I>MCS<SUB>3</SUB></I>>]
  edge []
  P0 -> E0, I0, C0, Y0, P1
  P1 -> E1, I1, C1, Y1, P2
  P2 -> E2, I2, C2, Y2
  E0 -> I0, C0, Y0, P1, E1
  E1 -> I1, C1, Y1, P2, E2
  E2 -> I2, C2, Y2
  I0 -> C0, Y0, P1, E1, I1
  I1 -> C1, Y1, P2, E2, I2
  I2 -> C2, Y2
  C0 -> Y0, P1, E1, I1, C1
  C1 -> Y1, P2, E2, I2, C2
  C2 -> Y2
  Y0 -> Y1 -> Y2
  Y0 -> E1
  Y1 -> E2
  {rank=min; P0, P1, P2}
  {rank=same; E0, E1, E2}
  {rank=same; I0, I1, I2}
  {rank=same; C0, C1, C2}
  {rank=max; Y0, Y1, Y2}
}")
```

## Setup wide data for MI

```{r setup-wide}
wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |> 
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12mcs_dv_base # wave 1
                  ), 
    outcome = sf12mcs_dv,
    pcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    econ_dist,
    static = FALSE,
    waves = c(1,2,3)
  )

if (file.exists(here::here("data", "wide_mids_3_waves_mcs_dynamic.rds"))) {
wide_mids <- readRDS(here::here("data", "wide_mids_3_waves_mcs_dynamic.rds"))
message("Loaded mice from 'data'")

} else {
if (!dir.exists(here::here("data")))
{
dir.create(here::here("data"))
message("Directory 'data' did not exist. Creating directory 'data'")
}
message("Creating wide_mids.rds into 'data'")
wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
saveRDS(wide_mids, file = here::here("data", "wide_mids_3_waves_mcs_dynamic.rds"))
message("Computed wide_mids and saved to 'data'")
}

#wide_mids <- readRDS(here::here("data", "wide_mids_full_models_MI_only.rds"))

base_predictor_matrix <- make_predictor_matrix(wide_data)

#information for descriptive table
regime_obs_n <- wide_data |>
  mutate(across(c(econ_emp_bin_1, econ_emp_bin_2, econ_emp_bin_3), as.character)) |>
  filter(!is.na(econ_emp_bin_1), !is.na(econ_emp_bin_2), !is.na(econ_emp_bin_3)) |>
  count(econ_emp_bin_1, econ_emp_bin_2, econ_emp_bin_3) |>
  mutate(intervention = paste(econ_emp_bin_1, econ_emp_bin_2, econ_emp_bin_3, sep = "-")) |>
  select(intervention, n_obs = n)
```

## gFormulaMI with accumulating variables: three waves of intervention

```{r 3waves_mi}
regimes <- intervention_pattern_3

predictor_matrix <- base_predictor_matrix

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS, only by their immediate previous value
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

# MCS in period t (not baseline) predicts employment and income in the period t+1
predictor_matrix["sf12mcs_dv_1",c("econ_emp_bin_2", "log_income_2")] <- 1
predictor_matrix["sf12mcs_dv_2",c("econ_emp_bin_3", "log_income_3")] <- 1

# Economic distress does not affect income or employment of the next period
predictor_matrix["econ_dist_1",c("econ_emp_bin_2", "log_income_2")] <- 0
predictor_matrix["econ_dist_2",c("econ_emp_bin_3", "log_income_3")] <- 0

# MCS in wave t predicts employment at time t+1
predictor_matrix["econ_emp_bin_2", "sf12mcs_dv_1"] <- 1
predictor_matrix["econ_emp_bin_3", "sf12mcs_dv_2"] <- 1

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_3cummulative <- syntheticPool(fits)

regimes_3 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

## Table

```{r 3wavesmi-tab}
#| output: true
mod_3cummulative <- mi_3cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_3)

mod_3cummulative |>
  knitr::kable(digits = 3)
```

## Graph

```{r 3wavesmi-graph}
#| output: true
mod_3cummulative |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```

# Exploratory analysis: Three periods of intervention - without intermediate outcomes

```{r model-baseline3-dag}
#| echo: false
#| output: true
grViz("digraph {
  graph []
  node [fontname=Arial]
  P0 [label=<PCS<SUB>1</SUB>>]
  P1 [label=<PCS<SUB>2</SUB>>]
  P2 [label=<PCS<SUB>3</SUB>>]
  E0 [label=<Empl<SUB>1</SUB>>]
  E1 [label=<Empl<SUB>2</SUB>>]
  E2 [label=<Empl<SUB>3</SUB>>]
  I0 [label=<Inc<SUB>1</SUB>>]
  I1 [label=<Inc<SUB>2</SUB>>]
  I2 [label=<Inc<SUB>3</SUB>>]
  C0 [label=<Conf<SUP>Int</SUP><SUB>1</SUB>>]
  C1 [label=<Conf<SUP>Int</SUP><SUB>2</SUB>>]
  C2 [label=<Conf<SUP>Int</SUP><SUB>3</SUB>>]
  Y0 [label=<<I>MCS<SUB>1</SUB></I>>]
  Y2 [label=<<I>MCS<SUB>3</SUB></I>>]
  edge []
  P0 -> E0, I0, C0, Y0, P1
  P1 -> E1, I1, C1, P2
  P2 -> E2, I2, C2, Y2
  E0 -> I0, C0, Y0, P1, E1
  E1 -> I1, C1, P2, E2
  E2 -> I2, C2, Y2
  I0 -> C0, Y0, P1, E1, I1
  I1 -> C1, P2, E2, I2
  I2 -> C2, Y2
  C0 -> Y0, P1, E1, I1, C1
  C1 -> P2, E2, I2, C2
  C2 -> Y2
  Y0 -> Y2
  Y0 -> E1
  {rank=min; P0, P1, P2}
  {rank=same; E0, E1, E2}
  {rank=same; I0, I1, I2}
  {rank=same; C0, C1, C2}
  {rank=max; Y0, Y2}
}")
```

## Setup wide data for MI

```{r setup-wide-static}
wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |> 
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12mcs_dv_base # wave 1
                  ), 
    outcome = sf12mcs_dv,
    pcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    econ_dist,
    static = TRUE,
    waves = c(1,2,3)
  )

if (file.exists(here::here("data", "wide_mids_3_waves_mcs_static.rds"))) {
wide_mids <- readRDS(here::here("data", "wide_mids_3_waves_mcs_static.rds"))
message("Loaded mice from 'data'")

} else {
if (!dir.exists(here::here("data")))
{
dir.create(here::here("data"))
message("Directory 'data' did not exist. Creating directory 'data'")
}
message("Creating wide_mids.rds into 'data'")
wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
saveRDS(wide_mids, file = here::here("data", "wide_mids_3_waves_mcs_static.rds"))
message("Computed wide_mids and saved to 'data'")
}

#wide_mids <- readRDS(here::here("data", "wide_mids_full_models_MI_only.rds"))

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

## gFormulaMI - three waves of intervention, without intermediate outcomes

```{r 3static-mi}
regimes <- intervention_pattern_3

predictor_matrix <- base_predictor_matrix

# predictor_matrix["sf12mcs_dv_0",] <- 0
# predictor_matrix["sf12mcs_dv_1",] <- 0
# predictor_matrix["sf12mcs_dv_2",] <- 0

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_3"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_1"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

# MCS in period t (not baseline) predicts employment and income in the period t+1
predictor_matrix["sf12mcs_dv_1",c("econ_emp_bin_2", "log_income_2")] <- 1

# Economic distress does not affect income or employment of the next period
predictor_matrix["econ_dist_1",c("econ_emp_bin_2", "log_income_2")] <- 0

# MCS in wave t predicts employment at time t+1
predictor_matrix["econ_emp_bin_2", "sf12mcs_dv_1"] <- 1

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_3static <- syntheticPool(fits)

regimes_3 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)

regime_obs_n <- wide_data |>
  mutate(across(c(econ_emp_bin_1, econ_emp_bin_2, econ_emp_bin_3), as.character)) |>
  filter(!is.na(econ_emp_bin_1), !is.na(econ_emp_bin_2), !is.na(econ_emp_bin_3)) |>
  count(econ_emp_bin_1, econ_emp_bin_2, econ_emp_bin_3) |>
  mutate(intervention = paste(econ_emp_bin_1, econ_emp_bin_2, econ_emp_bin_3, sep = "-")) |>
  select(intervention, n_obs = n)
```

## Table

```{r 3staticmi-tab}
#| output: true
mod_3static <- mi_3static |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_3)

mod_3static |>
  knitr::kable(digits = 3)
```

## Graph

```{r 3staticmi-graph}
#| output: true
mod_3static |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```

# Three waves of intervention - mediation analysis of MCS\_{2} via descomposition

We will estimate the randomised interventional effect of MCS\_{2} on MCS\_{3} as described in [VanderWeele, TJ and Tchetgen, ET 2014](https://biostats.bepress.com/harvardbiostat/paper168/) and [Lin, SH et al. 2017](https://doi.org/10.1097/EDE.0000000000000609). This effect is calculated by the generalised (parametric) mediational g-formula as follows:

$$
\begin{aligned}
rNDE &= Q(\bar{a}, \bar{a^*}) - Q(\bar{a^*}, \bar{a^*}) 
\\[1pt]
IDE &= Q(\bar{a}, \bar{a}) - Q(\bar{a}, \bar{a^*})
\end{aligned}
$$

$IDE$ represents the interventional direct effect of employment on MCS\_{3} not operating through MCS\_{2}, while $rNDE$ represents the total effect (TE) of employment on MCS\_{3}. The interventional indirect effect ($IIE$) is then calculated as the difference between the total and direct effects ($IIE = TE - IDE$).

Where:
$$
\begin{aligned}
Q(\bar{a_{1}}, \bar{a_{2}}) &= \sum_{\bar{m}}\sum_{\bar{l}(T-1)} E(Y\mid\bar{a},\bar{m},\bar{l},v)\prod_{t=1}^{T-1}P(l(t)\mid \overline{a_{1}(t-1)}, \overline{m(t-1)}, \overline{l(t-1)}) \\ 
&\quad \times \sum_{\bar{l'}}\prod_{t=0}^{T-1}P(m(t)\mid \overline{a_{2}(t)}, \overline{m(t-1)}, \overline{l'(t)})P(l'(t)\mid\overline{a_{2}(t-1)}\overline{m(t-1)},\overline{l(t-1)})
\end{aligned}
$$

$G_{a*|C}$ represents a random draw from the distribution of MCS\_{2} under the reference regime (employment = 0) conditional on confounders. By "randomising" the mediator to draws from the reference-regime distribution in both arms, we break the Empl→MCS₂→MCS₃ pathway and isolate the direct effect of employment on MCS\_{3} not operating through MCS\_{2}. \

Identification requires the following assumptions:

- (i)   No unmeasured Empl–MCS₃ confounding given C
- (ii*) No unmeasured MCS₂–MCS₃ confounding given Empl, C, L
- (iii) No unmeasured Empl–MCS₂ confounding given C

Does NOT require the cross-world independence assumption needed for NDE/NIE. \

Identified even with the MCS\_t → Empl\_{t+1} feedback present in this DAG. \


```{r decomposition-setup}
#| output: false

# Identification (Approach 3 from VanderWeele et al. 2014):
#   (i)   No unmeasured Empl–MCS₃ confounding given C
#   (ii*) No unmeasured MCS₂–MCS₃ confounding given Empl, C, L
#   (iii) No unmeasured Empl–MCS₂ confounding given C
#   → Does NOT require the cross-world independence assumption needed for NDE/NIE.
#   → Identified even with the MCS_t → Empl_{t+1} feedback present in this DAG.
# ══════════════════════════════════════════════════════════════════════════════


# ── STEP 0: Reload the dynamic MICE object ────────────────────────────────────
# wide_mids was overwritten in the static-model section above.
# The RIE requires the dynamic model (sf12mcs_dv_2 must be present in the data).
wide_mids_dyn <- readRDS(here::here("data", "wide_mids_3_waves_mcs_dynamic.rds"))

# wide_mids_dyn$data is the original incomplete wide data passed to mice().
# The n_tvars attribute set by make_wide() is preserved in the RDS object.
wide_data_dyn <- wide_mids_dyn$data

# Verify: the n_tvars attribute must be present for make_predictor_matrix()
stopifnot(!is.null(attr(wide_data_dyn, "n_tvars")))


# ── STEP 1: Rebuild the predictor matrix ──────────────────────────────────────
# Exact copy of the modifications in the dynamic model chunk (3waves_mi).
# This ensures the MCS₂ and MCS₃ model specifications are identical.
pm_rie <- make_predictor_matrix(wide_data_dyn)

# MCS intermediate outcomes do not act as predictors of other intermediates
pm_rie[, "sf12mcs_dv_1"] <- 0
pm_rie[, "sf12mcs_dv_2"] <- 0

# MCS₂ is predicted by its immediate predecessor MCS₁ (not by baseline MCS)
pm_rie["sf12mcs_dv_2", "sf12mcs_dv_1"]   <- 1
pm_rie["sf12mcs_dv_2", "sf12mcs_dv_base"] <- 0

# MCS₃ (outcome) is predicted by MCS₂ (not by baseline MCS)
pm_rie["sf12mcs_dv_3", "sf12mcs_dv_2"]   <- 1
pm_rie["sf12mcs_dv_3", "sf12mcs_dv_base"] <- 0

# MCS₁ and MCS₂ predict employment and income at the next wave (feedback arrows)
pm_rie["sf12mcs_dv_1", c("econ_emp_bin_2", "log_income_2")] <- 1
pm_rie["sf12mcs_dv_2", c("econ_emp_bin_3", "log_income_3")] <- 1

# Economic distress does not carry forward to next-period employment or income
pm_rie["econ_dist_1", c("econ_emp_bin_2", "log_income_2")] <- 0
pm_rie["econ_dist_2", c("econ_emp_bin_3", "log_income_3")] <- 0

# MCS at time t is used as a predictor when imputing employment at t+1 (feedback)
pm_rie["econ_emp_bin_2", "sf12mcs_dv_1"] <- 1
pm_rie["econ_emp_bin_3", "sf12mcs_dv_2"] <- 1


# ── STEP 2: Extract model formulas from the predictor matrix ──────────────────
# Convention: pm_rie[i, j] = 1  →  variable j predicts variable i.
# "regime" is a column in the matrix but only exists in synthetic data; exclude it.

# Model for MCS₂ (intermediate mediator): predictors at t=1 and baseline
mcs2_preds   <- setdiff(names(which(pm_rie["sf12mcs_dv_2", ] == 1)), "regime")
mcs2_formula <- reformulate(mcs2_preds, response = "sf12mcs_dv_2")

# Model for MCS₃ (final outcome): predictors include MCS₂ — we will intervene on this
mcs3_preds   <- setdiff(names(which(pm_rie["sf12mcs_dv_3", ] == 1)), "regime")
mcs3_formula <- reformulate(mcs3_preds, response = "sf12mcs_dv_3")

# Verify the key variables are in the right models before running anything
stopifnot("sf12mcs_dv_1"   %in% mcs2_preds)   # MCS₁ → MCS₂
stopifnot("sf12mcs_dv_2"   %in% mcs3_preds)   # MCS₂ → MCS₃ (will be frozen in RIE)
stopifnot("econ_emp_bin_2" %in% mcs2_preds)   # Empl₂ → MCS₂ (the path we intervene on)
stopifnot("econ_emp_bin_3" %in% mcs3_preds)   # Empl₃ → MCS₃


# ── STEP 3: Two-stage gFormulaImpute ──────────────────────────────────────────
# We stay within the gFormulaMI framework throughout so that:
#   (a) the same MICE models drive all imputations (no re-fitting),
#   (b) intermediate confounders (PCS₂, Inc₂, …) are re-simulated sequentially
#       under the correct employment regime, and
#   (c) syntheticPool is used for variance estimation (consistent with the rest
#       of the file).
#
# The two stages are:
#
#   STAGE 1 — reference regime only, MCS₂ free to re-impute
#     gFormulaImpute sets employment = (0,0,0) and lets MCS₂ be re-drawn from
#     its MICE model given those reference employment values.
#     The resulting sf12mcs_dv_2 in each completed dataset IS G_{a*|C}.
#
#   STAGE 2 — all 8 regimes; then MCS₂ frozen to Stage 1 draws per imputation
#     gFormulaImpute re-simulates all intermediate confounders under each
#     employment regime. We extract each imputation's completed data,
#     replace sf12mcs_dv_2 with G_{a*|C} from Stage 1, and fit the outcome
#     model. Results are collected for syntheticPool.
#
# Key note: gFormulaImpute returns a mids object. Variable columns are NOT
# directly accessible via $ (e.g. ref_imps$sf12mcs_dv_2 returns NULL —
# that would look up a mids list slot, not a data column). We must use
# mice::complete(mids, action = k) to extract the k-th completed data frame.

# ── Stage 1: generate G_{a*|C} under reference employment ────────────────────
# trtRegimes must be a plain numeric vector (not list) for a single regime.
# gFormulaImpute's numRegimes==1 path uses trtRegimes[i] (single bracket).
# On a list, [i] returns a sub-list → assignment to data frame fails.
# A plain vector makes [i] return a scalar, which assigns correctly.
ref_imps <- gFormulaImpute(
  wide_mids_dyn,
  trtVars     = c("econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes  = c(0, 0, 0),   # plain vector: always employed (reference)
  predictorMatrix = pm_rie
)

# ── Stage 2: all 8 regimes ────────────────────────────────────────────────────
rie_imps <- gFormulaImpute(
  wide_mids_dyn,
  trtVars     = c("econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes  = intervention_pattern_3,
  predictorMatrix = pm_rie
)

# ── Merge + fit: freeze MCS₂ to G_{a*|C} per imputation ─────────────────────
# For each synthetic imputation k:
#   ref_k  = mice::complete(ref_imps, k) → nSim rows (1 regime)
#   rie_k  = mice::complete(rie_imps, k) → nSim × 8 rows (8 regime blocks)
# We tile ref_k$sf12mcs_dv_2 across the 8 blocks in rie_k, then fit lm.
# syntheticPool expects a list with element $analyses (list of lm fits).
M <- wide_mids_dyn$m

# Sanity check: verify nSim and regime count from the mids objects
n_ref_per_imp <- nrow(mice::complete(ref_imps, action = 1))
n_rie_per_imp <- nrow(mice::complete(rie_imps, action = 1))
n_regimes     <- n_rie_per_imp / n_ref_per_imp
stopifnot(n_regimes == length(intervention_pattern_3))   # should equal 8

fits_analyses <- lapply(seq_len(M), function(k) {
  ref_k <- mice::complete(ref_imps, action = k)   # reference-regime synthetic data
  rie_k <- mice::complete(rie_imps, action = k)   # all-regime synthetic data
  # Freeze MCS₂: tile G_{a*|C} from ref_k across the 8 regime blocks in rie_k
  rie_k$sf12mcs_dv_2 <- rep(ref_k$sf12mcs_dv_2, times = n_regimes)
  lm(sf12mcs_dv_3 ~ 0 + factor(regime), data = rie_k)
})

# ── STEP 4: Pool with syntheticPool ───────────────────────────────────────────
# syntheticPool accesses fits$analyses (list of lm objects), consistent with
# the mira objects produced by %$% in the dynamic and static model chunks.
fits_rie <- list(analyses = fits_analyses)

mi_rie <- syntheticPool(fits_rie)

rie_pooled <- mi_rie |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(
    rie_effect = Estimate,
    rie_se     = sqrt(Total),
    rie_ll     = `95% CI L`,
    rie_ul     = `95% CI U`
  ) |>
  bind_cols(regimes_3)


# ── STEP 5: Compute IDE and IIE ───────────────────────────────────────────────
# All effects are contrasted against the reference regime: always employed (0-0-0).
#
# IDE = E[MCS₃(Empl=a, G_{a*|C})] − E[MCS₃(Empl=a*, G_{a*|C})]
#     = rie_pooled[regime=a] − rie_pooled[regime="0-0-0"]
#     (both arms use the same frozen reference MCS₂ distribution)
#
# IIE = TE − IDE
#     = (mod_3cummulative[regime=a] − mod_3cummulative[regime="0-0-0"]) − IDE
#     (TE comes from the dynamic model where MCS₂ is free to respond to employment)

ref_label <- "0-0-0"

# Extract reference-regime RIE mean (denominator of IDE).
# rie_pooled has an "intervention" column (added via bind_cols(regimes_3) above).
rie_ref_mean <- rie_pooled$rie_effect[rie_pooled$intervention == ref_label]

# Extract reference-regime TE from the dynamic model.
# mod_3cummulative is already a processed tibble with mi_effect + intervention columns
# (created in chunk 3wavesmi-tab). No need to re-apply as_tibble()/bind_cols().
te_ref_mean <- mod_3cummulative |>
  filter(intervention == ref_label) |>
  pull(mi_effect)

rie_decomposition <- rie_pooled |>
  left_join(
    mod_3cummulative |>
      transmute(intervention, te_mean = mi_effect),
    by = "intervention"
  ) |>
  mutate(
    # Total effect: contrast vs reference regime
    TE  = te_mean - te_ref_mean,
    # Interventional Direct Effect: MCS₂ frozen, employment varies
    IDE = rie_effect - rie_ref_mean,
    # Interventional Indirect Effect: remainder = effect operating through MCS₂
    IIE = TE - IDE
  ) |>
  select(intervention, TE, IDE, IIE, rie_effect, rie_se, rie_ll, rie_ul)
```

## Table
```{r decomposition-table}
#| output: true

rie_decomposition |>
  knitr::kable(digits = 3)
```