---
title: "Full models - 5 waves, new imputation"
author: "Darwin Del Castillo"
format: 
  html:
    toc: true
    fig-format: svg
execute: 
  output: false
---
# New plan

- Check if with more datasets (m) we get more stable estimates
  - Previusuly used m = 10
  - Now use m = 50  
- Check if the iterations improve something
  - Previously used maxit = 10
  - Now use maxit = 20 -> Now we are here
- Try introducing more waves
  - First try with 4 waves is in file 08
  - File 11 will introduce five waves
  - Try to capture years between 2015-2019 (check documentation)
- Verify predictor matrix in MI method and g-formula
- Additional: improve make_wide function. 
  - Currently, seems fragile in some extreme situations
  - Works well, but try to make it more robust

Combine via Rubin's rules: (review if this is correct, compare with Bartlett et al. 2025)

Combined mean:

$$
\bar{\theta} = \frac{1}{m}\left (\sum_{i=1}^m{\theta_i}\right )
$$

Within-imputation pooled variance:

$$
V_W = \frac{1}{m} \sum_{i=1}^m{SE_i^2}
$$

Between-imputation variance:

$$
V_B = \frac{\sum_{i=1}^m (\theta_i - \overline{\theta})^2}{m-1}
$$

Combined variance:

$$
V_{Total} = V_W + V_B + \frac{V_B}{m}
$$

# Part 1: four periods of intervention, mice with m = 35 vs m = 50

> Note: all these models assumes `mcs` as time-varying confounder (dynamic models)

## Theory (add DAGs later) 

### Confounders

Baseline static confounders:

 - Sex
 - Age at baseline (and age squared)
 - Ethnicity
 - Education
 - Location (`gor_dv`)
 - Relationship status (`mastat_dv`)
 
Time-varying confounders:

 - Mental health (in updating models; `sf12mcs_dv`)
 - Physical health (`sf12pcs_dv`)
 - Financial distress (`econ_dist`)
 - Number of dependants (`dnc`)
 - Any income from benefits (`econ_benefits`)
 - Home ownership (`home_owner`)
 - Income (`log_income`)
 
Exposures:

 - Employment at waves 3, 4, 5, 6 (`econ_emp_bin`)


## Setup
```{r setup}
#| output: false
here::i_am("building_models/11_new_full_models.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gFormulaMI)
library(magrittr)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(mice)

pop_data <- import_data() |> 
  select(-econ_incquint) 


ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 2 (baseline), and wave 6

pop_dt <- pop_data |> 
  filter(wave == 2, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(wave == 6, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(!is.na(mastat_dv_base)) |> 
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |> 
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 2,
  pre_mcs = shift(sf12mcs_dv, type = "lag"),
  pre_pcs = shift(sf12pcs_dv, type = "lag")
)][wave %in% 3:6]

no_mh_outcomes <- long_data |> 
  group_by(pidp) |> 
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |> 
  filter(n_missing == 4, base_missing) |> 
  select(pidp) |> 
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)] 

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```


## Imputation with m = 35
```{r impute-two-datasets}
wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |> 
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  wnw_race_base,
                  hiqual_dv_base,
                  age_dv_base,
                  gor_dv_base,
                  mastat_dv_base,
                  sf12mcs_dv_base),
    outcome = sf12mcs_dv,
    pre_mcs,
    pre_pcs,
    econ_dist,
    dnc,
    econ_benefits,
    home_owner,
    econ_emp_bin,
    log_income
)

t0 <- Sys.time()
message("Imputing wide data, starting at ", t0)
wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)

message("Finished imputation at: ", format(Sys.time()),
        " | elapsed: ", round(difftime(Sys.time(), t0, units="mins"), 2), " min")
rm(t0)

base_predictor_matrix <- make_predictor_matrix(wide_data)

```

### Diagnostics: convergence

```{r 4dynamic-diag} 
#| output: true
plot(wide_mids, c("econ_emp_bin_1", "pre_mcs_1", "pre_pcs_1",
                     "econ_emp_bin_2", "pre_mcs_2", "pre_pcs_2",
                     "econ_emp_bin_3", "sf12mcs_dv_3", "pre_mcs_3", "pre_pcs_3"))
```

### Diagnostics: distributions
```{r 4dynamic-diag-2}
#| output: true
densityplot(wide_mids, ~ pre_mcs_1 + pre_mcs_2 + pre_mcs_3 + sf12mcs_dv_3)

densityplot(wide_mids, ~ pre_pcs_1 + pre_pcs_2 + pre_pcs_3)

densityplot(wide_mids, ~ econ_emp_bin_1 + econ_emp_bin_2 + econ_emp_bin_3)
```

## gFormulaMI

```{r 4dynamic-mi}
t0 <- Sys.time()
message("Running g-formula via MI, starting at ", t0)

regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

## predictor_matrix["sf12mcs_dv_base",] <- 0
## predictor_matrix[,"sf12mcs_dv_base"] <- 0

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_0"] <- 0
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_0","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4dynamic <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)

message("Finished g-computation via MI at: ", format(Sys.time()),
        " | elapsed: ", round(difftime(Sys.time(), t0, units="mins"), 2), " min")
rm(t0)
```

### Comparing results

```{r 4dynamic-tab}
#| output: true
mod_4dynamic <- mi_4dynamic |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

mod_4dynamic |> 
  knitr::kable(digits = 3)
```

```{r 4dynamic-graph}
#| output: true
mod_4dynamic |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```

## Imputation with m = 50
```{r impute-two-datasets-50}
t0 <- Sys.time()
message("Imputing wide data, starting at ", t0)
wide_mids_50 <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 50, maxit = 10)

message("Finished imputation at: ", format(Sys.time()),
        " | elapsed: ", round(difftime(Sys.time(), t0, units="mins"), 2), " min")
rm(t0)
```

### Diagnostics: convergence

```{r 4dynamic-diag-50} 
#| output: true
plot(wide_mids_50, c("econ_emp_bin_1", "pre_mcs_1", "pre_pcs_1",
                     "econ_emp_bin_2", "pre_mcs_2", "pre_pcs_2",
                     "econ_emp_bin_3", "sf12mcs_dv_3", "pre_mcs_3", "pre_pcs_3"))
```

### Diagnostics: distributions
```{r 4dynamic-diag-2-50}
#| output: true
densityplot(wide_mids, ~ pre_mcs_1 + pre_mcs_2 + pre_mcs_3 + sf12mcs_dv_3)

densityplot(wide_mids, ~ pre_pcs_1 + pre_pcs_2 + pre_pcs_3)

densityplot(wide_mids, ~ econ_emp_bin_1 + econ_emp_bin_2 + econ_emp_bin_3)
```

## gFormulaMI

```{r 4dynamic-mi-50}
t0 <- Sys.time()
message("Running g-formula via MI, starting at ", t0)

regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

## predictor_matrix["sf12mcs_dv_base",] <- 0
## predictor_matrix[,"sf12mcs_dv_base"] <- 0

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_0"] <- 0
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_0","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids_50,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4dynamic <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)

message("Finished g-computation via MI at: ", format(Sys.time()),
        " | elapsed: ", round(difftime(Sys.time(), t0, units="mins"), 2), " min")
rm(t0)
```

### Comparing results

```{r 4dynamic-tab-50}
#| output: true
mod_4dynamic <- mi_4dynamic |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

mod_4dynamic |> 
  knitr::kable(digits = 3)
```

```{r 4dynamic-graph-50}
#| output: true
mod_4dynamic |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```

## Imputation with m = 35 maxit = 20
```{r impute-two-datasets-35-20}
t0 <- Sys.time()
message("Imputing wide data, starting at ", t0)
wide_mids_35 <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 20)

message("Finished imputation at: ", format(Sys.time()),
        " | elapsed: ", round(difftime(Sys.time(), t0, units="mins"), 2), " min")
rm(t0)
```

### Diagnostics: convergence

```{r 4dynamic-diag-35-20} 
#| output: true
plot(wide_mids_35, c("econ_emp_bin_1", "pre_mcs_1", "pre_pcs_1",
                     "econ_emp_bin_2", "pre_mcs_2", "pre_pcs_2",
                     "econ_emp_bin_3", "sf12mcs_dv_3", "pre_mcs_3", "pre_pcs_3"))
```

### Diagnostics: distributions
```{r 4dynamic-diag-2-35-20}
#| output: true
densityplot(wide_mids, ~ pre_mcs_1 + pre_mcs_2 + pre_mcs_3 + sf12mcs_dv_3)

densityplot(wide_mids, ~ pre_pcs_1 + pre_pcs_2 + pre_pcs_3)

densityplot(wide_mids, ~ econ_emp_bin_1 + econ_emp_bin_2 + econ_emp_bin_3)
```

## gFormulaMI

```{r 4dynamic-mi-35-20}
t0 <- Sys.time()
message("Running g-formula via MI, starting at ", t0)

regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

## predictor_matrix["sf12mcs_dv_base",] <- 0
## predictor_matrix[,"sf12mcs_dv_base"] <- 0

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_0"] <- 0
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_0","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids_35,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4dynamic <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)

message("Finished g-computation via MI at: ", format(Sys.time()),
        " | elapsed: ", round(difftime(Sys.time(), t0, units="mins"), 2), " min")
rm(t0)
```

### Comparing results

```{r 4dynamic-tab-35-20}
#| output: true
mod_4dynamic <- mi_4dynamic |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

mod_4dynamic |> 
  knitr::kable(digits = 3)
```

```{r 4dynamic-graph-35-20}
#| output: true
mod_4dynamic |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1")
```