---
title: "Estimating four-period with baseline and accumulating variables"
author: "Darwin Del Castillo"
format: 
  html:
    toc: true
    fig-format: svg
execute: 
  output: false
---
# Four periods of intervention, analysed by blocks - accumulating treatment strategy

## DAG
### Dictionary

"sf12pcs_dv": Physical health \
"econ_emp_bin": Employment \
"log_income": Income \
"dnc": Number of children \ 
"home_owner": Home owner \
"econ_benefits": Receives any benefits \
"econ_dist": Economic distress \
"mastat_dv": Marital status \
"pre_mcs": Previous mental health score from waves 2,3,4,5 (not currently in use) \
"pre_pcs": Previous physical health score from waves 2,3,4,5 (not currently in use) \
"sf12mcs_dv": Mental health score \
"hiqual_dv_base": Highest qualification at baseline \
"wnw_race_base": Race (white/non-white) \

#### Legend:
- P0 \rightarrow PCS \
- E0 \rightarrow Employment \
- I0 \rightarrow Income \ 
- Y0 \rightarrow MCS \ 
- C0 \rightarrow Intermediate confounders \ 
- F0 \rightarrow Financial distress \
- X0 \rightarrow Age \
- Z \rightarrow Static confounders \

```{r model-dynamic4-dag-mcs}
#| echo: false
#| output: true

grViz("digraph {
  graph []
  node [fontname=Arial]
  X0 [label=<Age<SUB>t-1</SUB>>]
  X1 [label=<Age<SUB>t</SUB>>]
  C0 [label=<Int_Conf<SUB>t-1</SUB>>]
  C1 [label=<Int_Conf<SUB>t</SUB>>]
  E1 [label=<Empl<SUB>t</SUB>>]
  E2 [label=<Empl<SUB>t+1</SUB>>]
  I1 [label=<Inc<SUB>t</SUB>>]
  I2 [label=<Inc<SUB>t+1</SUB>>]
  F1 [label=<Fin_Dist<SUB>t</SUB>>]
  F2 [label=<Fin_Dist<SUB>t+1</SUB>>]
  Y0 [label=<<I>MCS<SUB>t-1</SUB></I>>]
  Y1 [label=<<I>MCS<SUB>t</SUB></I>>]
  Y2 [label=<<I>MCS<SUB>t+1</SUB></I>>]
  P0 [label=<PCS<SUB>t-1</SUB>>]
  P1 [label=<PCS<SUB>t</SUB>>]
  Z [label=<Static_Conf>]
  edge []
  X0 -> X1
  X0 -> C0, E1, Y1, I1, F1, P0
  X1 -> C1, E2, Y2, I2, F2, P1
  P0 -> P1
  P0 -> E1, I1, F1, Y1
  P1 -> E2, I2, F2, Y2
  E1 -> I1, C1, F1, Y1, P1
  E1 -> E2
  I1 -> F1, Y1, C1, P1, E2
  I1 -> I2
  I2 -> F2, Y2
  C0 -> C1
  C0 -> E1, I1, F1, Y1
  C1 -> E2, I2, F2, Y2
  F1 -> F2
  F1 -> C1, Y1, P1
  F2 -> Y2
  Y0 -> C0, I1, P0, E1
  Y0 -> Y1 -> Y2
  Z -> Y0, C0, P0, I1, F1, E1, Y1, P1, C1, E2, I2, F2, Y2
  {rank=min; Z}
  {rank=same; X0, X1}
  {rank=same; P0, P1}
  {rank=same; E1, E2}
  {rank=same; I1, I2}
  {rank=same; C0, C1}
  {rank=same; F1, F2}
  {rank=max; Y0, Y1, Y2}
}")
```

## Setup - wave 3-6, with baseline covariates and outcomes

```{r setup-wave3-6}
#| output: false
here::i_am("building_models/12_workflow_blocks.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(gFormulaMI)
library(magrittr)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(DiagrammeR)

pop_data <- import_data() |> 
  select(-econ_incquint) 

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 2 (baseline), and wave 6

pop_dt <- pop_data |> 
  filter(wave == 2, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(wave == 6, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(!is.na(mastat_dv_base)) |> 
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |> 
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 2,
#  pre_mcs = shift(sf12mcs_dv, type = "lag"),
#  pre_pcs = shift(sf12pcs_dv, type = "lag"),
  mcs_outcome = sf12mcs_dv,
  pcs_outcome = sf12pcs_dv
)][wave %in% 3:6]

no_mh_outcomes <- long_data |> 
  group_by(pidp) |> 
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |> 
  filter(n_missing == 2, base_missing) |> 
  select(pidp) |> 
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)] 

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave3-6}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |> 
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12mcs_dv_base # wave 1
                  ), 
    outcome = sf12mcs_dv, 
    sf12pcs_dv,
    econ_emp_bin,
    log_income,
    dnc,
    home_owner,
    econ_benefits,
    econ_dist,
    sf12mcs_dv
  )

wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
#wide_mids <- readRDS(here::here("data", "wide_mids_full_models_MI_only.rds"))

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave3-6}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# predictor_matrix |> as.data.frame() |>
# openxlsx::write.xlsx(here::here("data", "predictor_matrix.xlsx"), overwrite = TRUE, rowNames = TRUE, colNames = TRUE)

# predictor_matrix["sf12mcs_dv_0",] <- 0
# predictor_matrix["sf12mcs_dv_1",] <- 0
# predictor_matrix["sf12mcs_dv_2",] <- 0
# predictor_matrix["sf12mcs_dv_3",] <- 0

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_0"] <- 0
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0
predictor_matrix[,"sf12mcs_dv_3"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave3-6}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

mod_4cummulative |> 
  knitr::kable(digits = 3)
```

#### Graph with estimates

```{r 4cummulative-graph-wave3-6}
#| output: true
mod_4cummulative |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |> 
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

## Setup - wave 4-7, with baseline covariates and outcomes

```{r setup-wave4-7}
#| output: false
here::i_am("building_models/12_workflow_blocks.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(gFormulaMI)
library(magrittr)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(DiagrammeR)

pop_data <- import_data() |> 
  select(-econ_incquint) 

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 3 (baseline), and wave 7

pop_dt <- pop_data |> 
  filter(wave == 3, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(wave == 7, response == 1) |> 
  select(pidp) |> 
  left_join(pop_data, by = "pidp") |> 
  filter(!is.na(mastat_dv_base)) |> 
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |> 
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 3,
#  pre_mcs = shift(sf12mcs_dv, type = "lag"),
#  pre_pcs = shift(sf12pcs_dv, type = "lag"),
  mcs_outcome = sf12mcs_dv,
  pcs_outcome = sf12pcs_dv
)][wave %in% 4:7]

no_mh_outcomes <- long_data |> 
  group_by(pidp) |> 
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |> 
  filter(n_missing == 2, base_missing) |> 
  select(pidp) |> 
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)] 

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave4-7}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |> 
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12mcs_dv_base # wave 1
                  ), 
    outcome = sf12mcs_dv, 
    sf12pcs_dv,
    econ_emp_bin,
    log_income,
    dnc,
    home_owner,
    econ_benefits,
    econ_dist,
    sf12mcs_dv
  )

wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
#wide_mids <- readRDS(here::here("data", "wide_mids_full_models_MI_only.rds"))

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

```{r 4cummulative-mi-wave4-7}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# predictor_matrix |> as.data.frame() |>
# openxlsx::write.xlsx(here::here("data", "predictor_matrix.xlsx"), overwrite = TRUE, rowNames = TRUE, colNames = TRUE)

# predictor_matrix["sf12mcs_dv_0",] <- 0
# predictor_matrix["sf12mcs_dv_1",] <- 0
# predictor_matrix["sf12mcs_dv_2",] <- 0
# predictor_matrix["sf12mcs_dv_3",] <- 0

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_0"] <- 0
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0
predictor_matrix[,"sf12mcs_dv_3"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave4-7}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

mod_4cummulative |> 
  knitr::kable(digits = 3)
```

#### Graph with estimates

```{r 4cummulative-graph-wave4-7}
#| output: true
mod_4cummulative |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

## Setup - wave 5-8, with baseline covariates and outcomes

```{r setup-wave5-8}
#| output: false
here::i_am("building_models/12_workflow_blocks.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(gFormulaMI)
library(magrittr)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(DiagrammeR)

pop_data <- import_data() |>
  select(-econ_incquint)

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 4, and wave 8

pop_dt <- pop_data |>
  filter(wave == 4, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(wave == 8, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(!is.na(mastat_dv_base)) |>
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |>
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 4,
#  pre_mcs = shift(sf12mcs_dv, type = "lag"),
#  pre_pcs = shift(sf12pcs_dv, type = "lag"),
  mcs_outcome = sf12mcs_dv,
  pcs_outcome = sf12pcs_dv
)][wave %in% 5:8]

no_mh_outcomes <- long_data |>
  group_by(pidp) |>
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |>
  filter(n_missing == 2, base_missing) |>
  select(pidp) |>
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave5-8}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12mcs_dv_base # wave 1
                  ),
    outcome = sf12mcs_dv,
    sf12pcs_dv,
    econ_emp_bin,
    log_income,
    dnc,
    home_owner,
    econ_benefits,
    econ_dist,
    sf12mcs_dv
  )

wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave5-8}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_0"] <- 0
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0
predictor_matrix[,"sf12mcs_dv_3"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave5-8}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

mod_4cummulative |>
  knitr::kable(digits = 3)
```

#### Graph with estimates

```{r 4cummulative-graph-wave5-8}
#| output: true
mod_4cummulative |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

## Setup - wave 6-9, with baseline covariates and outcomes

```{r setup-wave6-9}
#| output: false
here::i_am("building_models/12_workflow_blocks.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(gFormulaMI)
library(magrittr)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(DiagrammeR)

pop_data <- import_data() |>
  select(-econ_incquint)

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 5, and wave 9

pop_dt <- pop_data |>
  filter(wave == 5, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(wave == 9, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(!is.na(mastat_dv_base)) |>
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |>
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 5,
#  pre_mcs = shift(sf12mcs_dv, type = "lag"),
#  pre_pcs = shift(sf12pcs_dv, type = "lag"),
  mcs_outcome = sf12mcs_dv,
  pcs_outcome = sf12pcs_dv
)][wave %in% 6:9]

no_mh_outcomes <- long_data |>
  group_by(pidp) |>
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |>
  filter(n_missing == 2, base_missing) |>
  select(pidp) |>
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave6-9}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12mcs_dv_base # wave 1
                  ),
    outcome = sf12mcs_dv,
    sf12pcs_dv,
    econ_emp_bin,
    log_income,
    dnc,
    home_owner,
    econ_benefits,
    econ_dist,
    sf12mcs_dv
  )

wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave6-9}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_0"] <- 0
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0
predictor_matrix[,"sf12mcs_dv_3"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave6-9}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

mod_4cummulative |>
  knitr::kable(digits = 3)
```

#### Graph with estimates

```{r 4cummulative-graph-wave6-9}
#| output: true
mod_4cummulative |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

## Setup - wave 7-10, with baseline covariates and outcomes

```{r setup-wave7-10}
#| output: false
here::i_am("building_models/12_workflow_blocks.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(gFormulaMI)
library(magrittr)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(DiagrammeR)

pop_data <- import_data() |>
  select(-econ_incquint)

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 6, and wave 10

pop_dt <- pop_data |>
  filter(wave == 6, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(wave == 10, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(!is.na(mastat_dv_base)) |>
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |>
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 6,
#  pre_mcs = shift(sf12mcs_dv, type = "lag"),
#  pre_pcs = shift(sf12pcs_dv, type = "lag"),
  mcs_outcome = sf12mcs_dv,
  pcs_outcome = sf12pcs_dv
)][wave %in% 7:10]

no_mh_outcomes <- long_data |>
  group_by(pidp) |>
  summarise(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))) |>
  filter(n_missing == 2, base_missing) |>
  select(pidp) |>
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave7-10}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12mcs_dv_base # wave 1
                  ),
    outcome = sf12mcs_dv,
    sf12pcs_dv,
    econ_emp_bin,
    log_income,
    dnc,
    home_owner,
    econ_benefits,
    econ_dist,
    sf12mcs_dv
  )

wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave7-10}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# MCS does not predict intermediate outcomes
predictor_matrix[,"sf12mcs_dv_0"] <- 0
predictor_matrix[,"sf12mcs_dv_1"] <- 0
predictor_matrix[,"sf12mcs_dv_2"] <- 0
predictor_matrix[,"sf12mcs_dv_3"] <- 0

# MCS is dependent on previous MCS values
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_1"] <- 1
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_2"] <- 1

# MCS is not predicted by baseline MCS
predictor_matrix["sf12mcs_dv_1","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_2","sf12mcs_dv_base"] <- 0
predictor_matrix["sf12mcs_dv_3","sf12mcs_dv_base"] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12mcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave7-10}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

mod_4cummulative |>
  knitr::kable(digits = 3)
```

#### Graph with estimates

```{r 4cummulative-graph-wave7-10}
#| output: true
mod_4cummulative |>
  pivot_longer(
    -intervention,
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
           factor()) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

