---
title: "Estimating four-period PCS with baseline and accumulating variables"
author: "Darwin Del Castillo"
format:
  html:
    toc: true
    fig-format: svg
execute:
  output: false
---
# Four periods of intervention, analysed by blocks - accumulating treatment strategy (PCS)

## DAG
### Dictionary
#### Time invariant confounders
"hiqual_dv_base": Highest qualification at baseline \
"wnw_race_base": Race (white/non-white) \
"sex_dv_base": Sex \
"age_dv_base": Age \
"gor_dv_base": Geographical region \

#### Time varying confounders
:::{.callout-important}
These are lagged variables, their information come from wave 2
:::
"sf12mcs_dv": Mental health \
"dnc": Number of children \
"home_owner": Home owner \
"econ_benefits": Receives any benefits \
"mastat_dv": Marital status \

#### Time varying outcome
"sf12pcs_dv": Physical health score \

#### Time varying mediators
"log_income": Income \
"econ_dist": Economic distress \

#### Time varying treatment
"econ_emp_bin": Binary variable which represents whether the individual experienced unemployment or not \

:::{.callout-note}
This DAG is a simplification of the protocol DAG, as it does not include all the waves of the study, only summarises them in three time periods. \
**Legacy variables:**\
"pre_mcs": Previous mental health score from waves 2,3,4,5 (not currently in use) \
"pre_pcs": Previous physical health score from waves 2,3,4,5 (not currently in use) \
:::

### Legend:
- P0 \rightarrow MCS \
- E0 \rightarrow Employment \
- I0 \rightarrow Income \
- Y0 \rightarrow PCS \
- C0 \rightarrow Intermediate confounders \
- F0 \rightarrow Financial distress \
- X0 \rightarrow Age \
- Z \rightarrow Static confounders \

```{r packages-setup}
here::i_am("building_models/13_pcs_blocks.qmd")
source(here::here("R", "data_import_and_cleaning.R"))
source(here::here("R", "func_prep_data.R"))

library(gfoRmula)
library(gFormulaMI)
library(magrittr)
library(SPHSUgraphs)
library(tidyverse)
library(data.table)
library(DiagrammeR)
```

```{r model-dynamic4-dag-pcs}
#| echo: false
#| output: true

grViz("digraph {
  graph []
  node [fontname=Arial]
  X0 [label=<Age<SUB>t-1</SUB>>]
  X1 [label=<Age<SUB>t</SUB>>]
  C0 [label=<Int_Conf<SUB>t-1</SUB>>]
  C1 [label=<Int_Conf<SUB>t</SUB>>]
  E1 [label=<Empl<SUB>t</SUB>>]
  E2 [label=<Empl<SUB>t+1</SUB>>]
  I1 [label=<Inc<SUB>t</SUB>>]
  I2 [label=<Inc<SUB>t+1</SUB>>]
  F1 [label=<Fin_Dist<SUB>t</SUB>>]
  F2 [label=<Fin_Dist<SUB>t+1</SUB>>]
  Y0 [label=<<I>PCS<SUB>t-1</SUB></I>>]
  Y1 [label=<<I>PCS<SUB>t</SUB></I>>]
  Y2 [label=<<I>PCS<SUB>t+1</SUB></I>>]
  P0 [label=<MCS<SUB>t-1</SUB>>]
  P1 [label=<MCS<SUB>t</SUB>>]
  Z [label=<Static_Conf>]
  edge []
  X0 -> X1
  X0 -> C0, E1, Y1, I1, F1, P0
  X1 -> C1, E2, Y2, I2, F2, P1
  P0 -> P1
  P0 -> E1, I1, F1, Y1
  P1 -> E2, I2, F2, Y2
  E1 -> I1, C1, F1, Y1, P1
  E1 -> E2
  I1 -> F1, Y1, C1, P1, E2
  I1 -> I2
  I2 -> F2, Y2
  C0 -> C1
  C0 -> E1, I1, F1, Y1
  C1 -> E2, I2, F2, Y2
  F1 -> F2
  F1 -> C1, Y1, P1
  F2 -> Y2
  Y0 -> C0, I1, P0, E1
  Y0 -> Y1 -> Y2
  Z -> Y0, C0, P0, I1, F1, E1, Y1, P1, C1, E2, I2, F2, Y2
  {rank=min; Z}
  {rank=same; X0, X1}
  {rank=same; P0, P1}
  {rank=same; E1, E2}
  {rank=same; I1, I2}
  {rank=same; C0, C1}
  {rank=same; F1, F2}
  {rank=max; Y0, Y1, Y2}
}")
```

## Setup - wave 3-6, with baseline covariates and outcomes

```{r setup-wave3-6}
#| output: false

pop_data <- import_data() |>
  select(-econ_incquint)

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 2 (baseline), and wave 6

pop_dt <- pop_data |>
  filter(wave == 2, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(wave == 6, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(!is.na(mastat_dv_base)) |>
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |>
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 2,
  mcs_lagged = shift(sf12mcs_dv, type = "lag"),
  lagged_dnc = shift(dnc, type = "lag"),
  lagged_home_owner = shift(home_owner, type = "lag"),
  lagged_econ_benefits = shift(econ_benefits, type = "lag"),
  pcs_outcome = sf12pcs_dv
)][wave %in% 3:6]

no_mh_outcomes <- long_data |>
  group_by(pidp) |>
  summarise(n_missing = sum(is.na(sf12pcs_dv)), base_missing = all(is.na(sf12pcs_dv_base))) |>
  filter(n_missing == 2, base_missing) |>
  select(pidp) |>
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave3-6}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12pcs_dv_base # wave 1
                  ),
    outcome = sf12pcs_dv,
    mcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    econ_dist,
    static = FALSE
  )

if (file.exists(here::here("data", "wide_mids_4_waves_pcs_dynamic_1.rds"))) {
wide_mids <- readRDS(here::here("data", "wide_mids_4_waves_pcs_dynamic_1.rds"))
message("Loaded mice from 'data'")

} else {
if (!dir.exists(here::here("data")))
{
dir.create(here::here("data"))
message("Directory 'data' did not exist. Creating directory 'data'")
}
message("Creating wide_mids.rds into 'data'")
wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
saveRDS(wide_mids, file = here::here("data", "wide_mids_4_waves_pcs_dynamic_1.rds"))
message("Computed wide_mids and saved to 'data'")
}

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave3-6}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# PCS does not predict intermediate outcomes
predictor_matrix[,"sf12pcs_dv_0"] <- 0
predictor_matrix[,"sf12pcs_dv_1"] <- 0
predictor_matrix[,"sf12pcs_dv_2"] <- 0
predictor_matrix[,"sf12pcs_dv_3"] <- 0

# PCS is dependent on previous PCS values
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_0"] <- 1
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_1"] <- 1
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_2"] <- 1

# PCS is not predicted by baseline PCS, only by their immediate previous value
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_base"] <- 0

# PCS in period t (not baseline) predicts employment and income in the period t+1
predictor_matrix["sf12pcs_dv_0",c("econ_emp_bin_1", "log_income_1")] <- 1
predictor_matrix["sf12pcs_dv_1",c("econ_emp_bin_2", "log_income_2")] <- 1
predictor_matrix["sf12pcs_dv_2",c("econ_emp_bin_3", "log_income_3")] <- 1

# Economic distress does not affect income or employment of the next period
predictor_matrix["econ_dist_0",c("econ_emp_bin_1", "log_income_1")] <- 0
predictor_matrix["econ_dist_1",c("econ_emp_bin_2", "log_income_2")] <- 0
predictor_matrix["econ_dist_2",c("econ_emp_bin_3", "log_income_3")] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12pcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave3-6}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

obs_counts <- long_data |>
  as_tibble() |>
  select(pidp, t0, econ_emp_bin) |>
  mutate(emp = as.integer(as.character(econ_emp_bin))) |>
  pivot_wider(id_cols = pidp, names_from = t0, values_from = emp, names_prefix = "period_") |>
  filter(!is.na(period_0), !is.na(period_1), !is.na(period_2), !is.na(period_3)) |>
  mutate(intervention = paste(period_0, period_1, period_2, period_3, sep = "-")) |>
  count(intervention, name = "n_obs")

mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  knitr::kable(digits = 3)

mod_4cummulative |> saveRDS(here::here("output", "mod_4cummulative_pcs_wave3-6.rds"))
```

#### Graph with estimates

```{r 4cummulative-graph-wave3-6}
#| output: true
mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  pivot_longer(
    -c(intervention, n_obs),
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |> factor(),
    intervention = paste0(intervention, " (n=", n_obs, ")")
  ) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

## Setup - wave 4-7, with baseline covariates and outcomes

```{r setup-wave4-7}
#| output: false

pop_data <- import_data() |>
  select(-econ_incquint)

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 3 (baseline), and wave 7

pop_dt <- pop_data |>
  filter(wave == 3, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(wave == 7, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(!is.na(mastat_dv_base)) |>
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |>
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 3,
  mcs_lagged = shift(sf12mcs_dv, type = "lag"),
  lagged_dnc = shift(dnc, type = "lag"),
  lagged_home_owner = shift(home_owner, type = "lag"),
  lagged_econ_benefits = shift(econ_benefits, type = "lag"),
  pcs_outcome = sf12pcs_dv
)][wave %in% 4:7]

no_mh_outcomes <- long_data |>
  group_by(pidp) |>
  summarise(n_missing = sum(is.na(sf12pcs_dv)), base_missing = all(is.na(sf12pcs_dv_base))) |>
  filter(n_missing == 2, base_missing) |>
  select(pidp) |>
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave4-7}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12pcs_dv_base # wave 1
                  ),
    outcome = sf12pcs_dv,
    mcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    econ_dist,
    static = FALSE
  )

if (file.exists(here::here("data", "wide_mids_4_waves_pcs_dynamic_2.rds"))) {
wide_mids <- readRDS(here::here("data", "wide_mids_4_waves_pcs_dynamic_2.rds"))
message("Loaded mice from 'data'")

} else {
if (!dir.exists(here::here("data")))
{
dir.create(here::here("data"))
message("Directory 'data' did not exist. Creating directory 'data'")
}
message("Creating wide_mids.rds into 'data'")
wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
saveRDS(wide_mids, file = here::here("data", "wide_mids_4_waves_pcs_dynamic_2.rds"))
message("Computed wide_mids and saved to 'data'")
}

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave4-7}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# PCS does not predict intermediate outcomes
predictor_matrix[,"sf12pcs_dv_0"] <- 0
predictor_matrix[,"sf12pcs_dv_1"] <- 0
predictor_matrix[,"sf12pcs_dv_2"] <- 0
predictor_matrix[,"sf12pcs_dv_3"] <- 0

# PCS is dependent on previous PCS values
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_0"] <- 1
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_1"] <- 1
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_2"] <- 1

# PCS is not predicted by baseline PCS
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_base"] <- 0

# PCS in period t (not baseline) predicts employment and income in the period t+1
predictor_matrix["sf12pcs_dv_0",c("econ_emp_bin_1", "log_income_1")] <- 1
predictor_matrix["sf12pcs_dv_1",c("econ_emp_bin_2", "log_income_2")] <- 1
predictor_matrix["sf12pcs_dv_2",c("econ_emp_bin_3", "log_income_3")] <- 1

# Economic distress does not affect income or employment of the next period
predictor_matrix["econ_dist_0",c("econ_emp_bin_1", "log_income_1")] <- 0
predictor_matrix["econ_dist_1",c("econ_emp_bin_2", "log_income_2")] <- 0
predictor_matrix["econ_dist_2",c("econ_emp_bin_3", "log_income_3")] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12pcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave4-7}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

obs_counts <- long_data |>
  as_tibble() |>
  select(pidp, t0, econ_emp_bin) |>
  mutate(emp = as.integer(as.character(econ_emp_bin))) |>
  pivot_wider(id_cols = pidp, names_from = t0, values_from = emp, names_prefix = "period_") |>
  filter(!is.na(period_0), !is.na(period_1), !is.na(period_2), !is.na(period_3)) |>
  mutate(intervention = paste(period_0, period_1, period_2, period_3, sep = "-")) |>
  count(intervention, name = "n_obs")

mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  knitr::kable(digits = 3)

mod_4cummulative |> saveRDS(here::here("output", "mod_4cummulative_pcs_wave4-7.rds"))
```

#### Graph with estimates

```{r 4cummulative-graph-wave4-7}
#| output: true
mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  pivot_longer(
    -c(intervention, n_obs),
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |> factor(),
    intervention = paste0(intervention, " (n=", n_obs, ")")
  ) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

## Setup - wave 5-8, with baseline covariates and outcomes

```{r setup-wave5-8}
#| output: false

pop_data <- import_data() |>
  select(-econ_incquint)

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 4, and wave 8

pop_dt <- pop_data |>
  filter(wave == 4, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(wave == 8, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(!is.na(mastat_dv_base)) |>
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |>
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 4,
  mcs_lagged = shift(sf12mcs_dv, type = "lag"),
  lagged_dnc = shift(dnc, type = "lag"),
  lagged_home_owner = shift(home_owner, type = "lag"),
  lagged_econ_benefits = shift(econ_benefits, type = "lag"),
  pcs_outcome = sf12pcs_dv
)][wave %in% 5:8]

no_mh_outcomes <- long_data |>
  group_by(pidp) |>
  summarise(n_missing = sum(is.na(sf12pcs_dv)), base_missing = all(is.na(sf12pcs_dv_base))) |>
  filter(n_missing == 2, base_missing) |>
  select(pidp) |>
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave5-8}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12pcs_dv_base # wave 1
                  ),
    outcome = sf12pcs_dv,
    mcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    econ_dist,
    static = FALSE
  )

if (file.exists(here::here("data", "wide_mids_4_waves_pcs_dynamic_3.rds"))) {
wide_mids <- readRDS(here::here("data", "wide_mids_4_waves_pcs_dynamic_3.rds"))
message("Loaded mice from 'data'")

} else {
if (!dir.exists(here::here("data")))
{
dir.create(here::here("data"))
message("Directory 'data' did not exist. Creating directory 'data'")
}
message("Creating wide_mids.rds into 'data'")
wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
saveRDS(wide_mids, file = here::here("data", "wide_mids_4_waves_pcs_dynamic_3.rds"))
message("Computed wide_mids and saved to 'data'")
}

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave5-8}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# PCS does not predict intermediate outcomes
predictor_matrix[,"sf12pcs_dv_0"] <- 0
predictor_matrix[,"sf12pcs_dv_1"] <- 0
predictor_matrix[,"sf12pcs_dv_2"] <- 0
predictor_matrix[,"sf12pcs_dv_3"] <- 0

# PCS is dependent on previous PCS values
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_0"] <- 1
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_1"] <- 1
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_2"] <- 1

# PCS is not predicted by baseline PCS
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_base"] <- 0

# PCS in period t (not baseline) predicts employment and income in the period t+1
predictor_matrix["sf12pcs_dv_0",c("econ_emp_bin_1", "log_income_1")] <- 1
predictor_matrix["sf12pcs_dv_1",c("econ_emp_bin_2", "log_income_2")] <- 1
predictor_matrix["sf12pcs_dv_2",c("econ_emp_bin_3", "log_income_3")] <- 1

# Economic distress does not affect income or employment of the next period
predictor_matrix["econ_dist_0",c("econ_emp_bin_1", "log_income_1")] <- 0
predictor_matrix["econ_dist_1",c("econ_emp_bin_2", "log_income_2")] <- 0
predictor_matrix["econ_dist_2",c("econ_emp_bin_3", "log_income_3")] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12pcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave5-8}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

obs_counts <- long_data |>
  as_tibble() |>
  select(pidp, t0, econ_emp_bin) |>
  mutate(emp = as.integer(as.character(econ_emp_bin))) |>
  pivot_wider(id_cols = pidp, names_from = t0, values_from = emp, names_prefix = "period_") |>
  filter(!is.na(period_0), !is.na(period_1), !is.na(period_2), !is.na(period_3)) |>
  mutate(intervention = paste(period_0, period_1, period_2, period_3, sep = "-")) |>
  count(intervention, name = "n_obs")

mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  knitr::kable(digits = 3)

mod_4cummulative |> saveRDS(here::here("output", "mod_4cummulative_pcs_wave5-8.rds"))
```

#### Graph with estimates

```{r 4cummulative-graph-wave5-8}
#| output: true
mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  pivot_longer(
    -c(intervention, n_obs),
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |> factor(),
    intervention = paste0(intervention, " (n=", n_obs, ")")
  ) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

## Setup - wave 6-9, with baseline covariates and outcomes

```{r setup-wave6-9}
#| output: false

pop_data <- import_data() |>
  select(-econ_incquint)

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 5, and wave 9

pop_dt <- pop_data |>
  filter(wave == 5, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(wave == 9, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(!is.na(mastat_dv_base)) |>
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |>
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 5,
  mcs_lagged = shift(sf12mcs_dv, type = "lag"),
  lagged_dnc = shift(dnc, type = "lag"),
  lagged_home_owner = shift(home_owner, type = "lag"),
  lagged_econ_benefits = shift(econ_benefits, type = "lag"),
  pcs_outcome = sf12pcs_dv
)][wave %in% 6:9]

no_mh_outcomes <- long_data |>
  group_by(pidp) |>
  summarise(n_missing = sum(is.na(sf12pcs_dv)), base_missing = all(is.na(sf12pcs_dv_base))) |>
  filter(n_missing == 2, base_missing) |>
  select(pidp) |>
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave6-9}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12pcs_dv_base # wave 1
                  ),
    outcome = sf12pcs_dv,
    mcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    econ_dist,
    static = FALSE
  )

if (file.exists(here::here("data", "wide_mids_4_waves_pcs_dynamic_4.rds"))) {
wide_mids <- readRDS(here::here("data", "wide_mids_4_waves_pcs_dynamic_4.rds"))
message("Loaded mice from 'data'")

} else {
if (!dir.exists(here::here("data")))
{
dir.create(here::here("data"))
message("Directory 'data' did not exist. Creating directory 'data'")
}
message("Creating wide_mids.rds into 'data'")
wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
saveRDS(wide_mids, file = here::here("data", "wide_mids_4_waves_pcs_dynamic_4.rds"))
message("Computed wide_mids and saved to 'data'")
}

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave6-9}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# PCS does not predict intermediate outcomes
predictor_matrix[,"sf12pcs_dv_0"] <- 0
predictor_matrix[,"sf12pcs_dv_1"] <- 0
predictor_matrix[,"sf12pcs_dv_2"] <- 0
predictor_matrix[,"sf12pcs_dv_3"] <- 0

# PCS is dependent on previous PCS values
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_0"] <- 1
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_1"] <- 1
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_2"] <- 1

# PCS is not predicted by baseline PCS
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_base"] <- 0

# PCS in period t (not baseline) predicts employment and income in the period t+1
predictor_matrix["sf12pcs_dv_0",c("econ_emp_bin_1", "log_income_1")] <- 1
predictor_matrix["sf12pcs_dv_1",c("econ_emp_bin_2", "log_income_2")] <- 1
predictor_matrix["sf12pcs_dv_2",c("econ_emp_bin_3", "log_income_3")] <- 1

# Economic distress does not affect income or employment of the next period
predictor_matrix["econ_dist_0",c("econ_emp_bin_1", "log_income_1")] <- 0
predictor_matrix["econ_dist_1",c("econ_emp_bin_2", "log_income_2")] <- 0
predictor_matrix["econ_dist_2",c("econ_emp_bin_3", "log_income_3")] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12pcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave6-9}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

obs_counts <- long_data |>
  as_tibble() |>
  select(pidp, t0, econ_emp_bin) |>
  mutate(emp = as.integer(as.character(econ_emp_bin))) |>
  pivot_wider(id_cols = pidp, names_from = t0, values_from = emp, names_prefix = "period_") |>
  filter(!is.na(period_0), !is.na(period_1), !is.na(period_2), !is.na(period_3)) |>
  mutate(intervention = paste(period_0, period_1, period_2, period_3, sep = "-")) |>
  count(intervention, name = "n_obs")

mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  knitr::kable(digits = 3)

mod_4cummulative |> saveRDS(here::here("output", "mod_4cummulative_pcs_wave6-9.rds"))
```

#### Graph with estimates

```{r 4cummulative-graph-wave6-9}
#| output: true
mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  pivot_longer(
    -c(intervention, n_obs),
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |> factor(),
    intervention = paste0(intervention, " (n=", n_obs, ")")
  ) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```

## Setup - wave 7-10, with baseline covariates and outcomes

```{r setup-wave7-10}
#| output: false

pop_data <- import_data() |>
  select(-econ_incquint)

ggplot2::theme_set(theme_sphsu_light())

# Filter for only those observed at wave 6, and wave 10

pop_dt <- pop_data |>
  filter(wave == 6, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(wave == 10, response == 1) |>
  select(pidp) |>
  left_join(pop_data, by = "pidp") |>
  filter(!is.na(mastat_dv_base)) |>
  mutate(dnc = factor(dnc) |> fct_collapse(zero = "0", one = "1", other_level = "2+")) |>
  data.table::as.data.table()

long_data <- pop_dt[, `:=`(
  t0 = t0 - 6,
  mcs_lagged = shift(sf12mcs_dv, type = "lag"),
  lagged_dnc = shift(dnc, type = "lag"),
  lagged_home_owner = shift(home_owner, type = "lag"),
  lagged_econ_benefits = shift(econ_benefits, type = "lag"),
  pcs_outcome = sf12pcs_dv
)][wave %in% 7:10]

no_mh_outcomes <- long_data |>
  group_by(pidp) |>
  summarise(n_missing = sum(is.na(sf12pcs_dv)), base_missing = all(is.na(sf12pcs_dv_base))) |>
  filter(n_missing == 2, base_missing) |>
  select(pidp) |>
  as.data.table()

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

intervention_pattern_4 <- expand_grid(0:1, 0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

ncores <- 2

nsamples <- 100
```

## Multiple imputation - setup of wide data for MI, and imputation model
```{r mids-setup-wave7-10}
#| output: false

# Setup wide data for MI

wide_data <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base, # wave 1
                  wnw_race_base, # wave 1
                  hiqual_dv_base, # wave 1
                  age_dv_base, # wave 1
                  gor_dv_base, # wave 1
                  mastat_dv_base, # wave 1
                  sf12pcs_dv_base # wave 1
                  ),
    outcome = sf12pcs_dv,
    mcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    econ_dist,
    static = FALSE
  )

if (file.exists(here::here("data", "wide_mids_4_waves_pcs_dynamic_5.rds"))) {
wide_mids <- readRDS(here::here("data", "wide_mids_4_waves_pcs_dynamic_5.rds"))
message("Loaded mice from 'data'")

} else {
if (!dir.exists(here::here("data")))
{
dir.create(here::here("data"))
message("Directory 'data' did not exist. Creating directory 'data'")
}
message("Creating wide_mids.rds into 'data'")
wide_mids <- mice::mice(wide_data, defaultMethod = c("norm", "logreg", "pmm", "pmm"), m = 35, maxit = 10)
saveRDS(wide_mids, file = here::here("data", "wide_mids_4_waves_pcs_dynamic_5.rds"))
message("Computed wide_mids and saved to 'data'")
}

base_predictor_matrix <- make_predictor_matrix(wide_data)
```

### gFormulaMI - setup of predictor matrix and regimes

```{r 4cummulative-mi-wave7-10}
regimes <- intervention_pattern_4

predictor_matrix <- base_predictor_matrix

# PCS does not predict intermediate outcomes
predictor_matrix[,"sf12pcs_dv_0"] <- 0
predictor_matrix[,"sf12pcs_dv_1"] <- 0
predictor_matrix[,"sf12pcs_dv_2"] <- 0
predictor_matrix[,"sf12pcs_dv_3"] <- 0

# PCS is dependent on previous PCS values
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_0"] <- 1
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_1"] <- 1
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_2"] <- 1

# PCS is not predicted by baseline PCS
predictor_matrix["sf12pcs_dv_1","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_2","sf12pcs_dv_base"] <- 0
predictor_matrix["sf12pcs_dv_3","sf12pcs_dv_base"] <- 0

# PCS in period t (not baseline) predicts employment and income in the period t+1
predictor_matrix["sf12pcs_dv_0",c("econ_emp_bin_1", "log_income_1")] <- 1
predictor_matrix["sf12pcs_dv_1",c("econ_emp_bin_2", "log_income_2")] <- 1
predictor_matrix["sf12pcs_dv_2",c("econ_emp_bin_3", "log_income_3")] <- 1

# Economic distress does not affect income or employment of the next period
predictor_matrix["econ_dist_0",c("econ_emp_bin_1", "log_income_1")] <- 0
predictor_matrix["econ_dist_1",c("econ_emp_bin_2", "log_income_2")] <- 0
predictor_matrix["econ_dist_2",c("econ_emp_bin_3", "log_income_3")] <- 0

imps <- gFormulaImpute(
  wide_mids,
  trtVars = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2", "econ_emp_bin_3"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix
)

fits <- imps %$%
  lm(sf12pcs_dv_3 ~ 0 + factor(regime))

mi_4cummulative <- syntheticPool(fits)

regimes_4 <- tibble(intervention = regimes |>
  map(paste, collapse = "-") |>
  reduce(c)
)
```

#### Table with estimates

```{r 4cummulative-tab-wave7-10}
#| output: true

mod_4cummulative <- mi_4cummulative |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_4)

obs_counts <- long_data |>
  as_tibble() |>
  select(pidp, t0, econ_emp_bin) |>
  mutate(emp = as.integer(as.character(econ_emp_bin))) |>
  pivot_wider(id_cols = pidp, names_from = t0, values_from = emp, names_prefix = "period_") |>
  filter(!is.na(period_0), !is.na(period_1), !is.na(period_2), !is.na(period_3)) |>
  mutate(intervention = paste(period_0, period_1, period_2, period_3, sep = "-")) |>
  count(intervention, name = "n_obs")

mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  knitr::kable(digits = 3)

mod_4cummulative |> saveRDS(here::here("output", "mod_4cummulative_pcs_wave7-10.rds"))
```

#### Graph with estimates

```{r 4cummulative-graph-wave7-10}
#| output: true
mod_4cummulative |>
  left_join(obs_counts, by = "intervention") |>
  replace_na(list(n_obs = 0)) |>
  pivot_longer(
    -c(intervention, n_obs),
    names_to = c("Calculation method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    n_unemp = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |> factor(),
    intervention = paste0(intervention, " (n=", n_obs, ")")
  ) |>
ggplot(aes(effect, intervention, shape = `Calculation method`, xmin = ll, xmax = ul,
    colour = n_unemp)) +
  geom_point(size = 2, position = position_dodge(width = 0.7))  +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.7)) +
  scale_colour_brewer(name = "Number of unemployed periods", palette = "Set1") +
  facet_wrap(~`Calculation method`, ncol = 1)
```
