---
title: "Effect Modification by Sex, Age Group, and Education"
author: "Darwin Del Castillo"
format:
  html:
    toc: true
    fig-format: svg
execute:
  output: false
---

```{r env-setup}
#| output: false
#| message: false
options(max.print = 999999)
here::i_am("building_models/20_effect_mod.qmd")
source(here::here("R", "import_and_clean.R"))
source(here::here("R", "func_prep_data.R"))
source(here::here("R", "helpers.R"))

pacman::p_load(data.table, here, mice, gFormulaMI, magrittr, tidyverse,
               SPHSUgraphs, colorBlindness, patchwork, cowplot)

seed <- 20260410
```

# Data Preparation

Same eligibility criteria as `17_four_waves_final.qmd` (waves 3–6, observed at waves 2 and 6).
Three effect modifiers are created from baseline (wave 1) variables:

- **Sex**: `sex_dv_base` (Female / Male)
- **Age group**: `age_dv_base` dichotomised at 40 (Younger 25–40 / Older 41–65)
- **Education**: `hiqual_dv_base` grouped as Higher (codes 1–2) vs Lower (codes 3–5, 9)

```{r prep-data}
#| message: false
#| warning: false

pop_data <- clean_data() |> preproc_data()

eligible_pidp <- intersect(
  pop_data[wave == 2 & response == 1, pidp],
  pop_data[wave == 5 & response == 1, pidp]
)

pop_DT <- pop_data[pidp %in% eligible_pidp]

long_data <- pop_DT[, `:=`(
  t0                   = t0 - 2L,
  pcs_lagged           = shift(sf12pcs_dv,    type = "lag"),
  mcs_lagged           = shift(sf12mcs_dv,    type = "lag"),
  lagged_dnc           = shift(dnc,           type = "lag"),
  lagged_home_owner    = shift(home_owner,    type = "lag"),
  lagged_econ_benefits = shift(econ_benefits, type = "lag"),
  lagged_gor_dv_fact   = shift(gor_dv_fact,   type = "lag"),
  lagged_age_dv        = shift(age_dv,        type = "lag"),
  lagged_mastat_dv     = shift(mastat_dv,     type = "lag")
)][wave %in% 3:5]

no_mh_outcomes <- long_data[,
  .(n_missing   = sum(is.na(sf12mcs_dv)),
    base_missing = all(is.na(sf12mcs_dv_base))),
  by = pidp
][n_missing >= 2 & base_missing == TRUE, .(pidp)]

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

# --- Strata variables (derived from baseline, time-invariant across all rows) ---
long_data[, age_group := factor(
  fifelse(age_dv_base <= 40, "Younger (25-40)", "Older (41-65)")
)]

long_data[, edu_group := factor(fcase(
  hiqual_dv_base %in% c(1L, 2L),          "Higher",
  hiqual_dv_base %in% c(3L, 4L, 5L, 9L), "Lower",
  default = NA_character_
))]

intervention_pattern_3 <- asplit(as.matrix(CJ(0:1, 0:1, 0:1)), 1)

regimes_3 <- tibble(
  intervention = map_chr(intervention_pattern_3, ~ paste(.x, collapse = "-"))
)
```

# MCS Outcome

## Wide format

`age_dv_base`, `age_group`, and `edu_group` are included in `base_cols` so `make_predictor_matrix()` correctly identifies them in the base-variable block.

```{r wide-mcs}
wide_data_mcs <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  race_base,
                  hiqual_dv_base,
                  age_dv_base,   # needed for age_group filtering
                  age_group,     # strata: Younger / Older
                  edu_group,     # strata: Higher / Lower
                  sf12mcs_dv_base),
    outcome = sf12mcs_dv,
    pcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    lagged_mastat_dv,
    lagged_age_dv,
    lagged_gor_dv_fact,
    econ_dist,
    static = FALSE,
    waves  = c(0, 1, 2)
  )
```

## Multiple imputation

Strata variables and `age_dv_base` have their rows *and* columns zeroed in the MICE predictor matrix: they are not imputed and are excluded as predictors to avoid collinearity with `lagged_age_dv` and `hiqual_dv_base`.

```{r mice-mcs}
pred_mat_mcs <- make.predictorMatrix(wide_data_mcs)

# Standard collinearity exclusions
pred_mat_mcs[, c("lagged_age_dv_0",   "lagged_age_dv_1",   "lagged_age_dv_2")]   <- 0
pred_mat_mcs[, c("lagged_gor_dv_fact_0", "lagged_gor_dv_fact_1", "lagged_gor_dv_fact_2")] <- 0

# Strata / classification variables: don't impute, don't use as predictors
for (v in c("age_dv_base", "age_group", "edu_group")) {
  pred_mat_mcs[v, ] <- 0
  pred_mat_mcs[, v] <- 0
}

if (file.exists(here::here("output", "wide_mids_mcs_em.rds"))) {
  wide_mids_mcs <- readRDS(here::here("output", "wide_mids_mcs_em.rds"))
  message("Loading cached MCS MIDS (effect modification)...")
} else {
  message("Running MICE for MCS effect modification analysis...")
  wide_mids_mcs <- mice(
    wide_data_mcs,
    predictorMatrix = pred_mat_mcs,
    defaultMethod   = c("norm", "logreg", "pmm", "pmm"),
    m               = 50,
    maxit           = 15,
    seed            = seed
  )
  saveRDS(wide_mids_mcs, here::here("output", "wide_mids_mcs_em.rds"))
}
```

## gFormulaMI predictor matrix

Same DAG structure as `16_three_waves_final.qmd`, plus strata variable columns
zeroed to prevent collinearity in the sequential imputation.

```{r pred-matrix-mcs}
predictor_matrix_mcs <- make_predictor_matrix(wide_data_mcs)

# Zero strata and collinear columns
predictor_matrix_mcs[, c("age_dv_base", "age_group", "edu_group")] <- 0

# MCS intermediate values do not predict other intermediates (block forward contamination)
predictor_matrix_mcs[, "sf12mcs_dv_0"] <- 0
predictor_matrix_mcs[, "sf12mcs_dv_1"] <- 0

# MCS predicted by its own immediate lag, not by baseline MCS
predictor_matrix_mcs["sf12mcs_dv_1", "sf12mcs_dv_0"]    <- 1
predictor_matrix_mcs["sf12mcs_dv_2", "sf12mcs_dv_1"]    <- 1
predictor_matrix_mcs["sf12mcs_dv_1", "sf12mcs_dv_base"] <- 0
predictor_matrix_mcs["sf12mcs_dv_2", "sf12mcs_dv_base"] <- 0

# MCS at t predicts employment and income at t+1 (treatment-confounder feedback)
predictor_matrix_mcs["sf12mcs_dv_0", c("econ_emp_bin_1", "log_income_1")] <- 1
predictor_matrix_mcs["sf12mcs_dv_1", c("econ_emp_bin_2", "log_income_2")] <- 1

# Economic distress does not propagate to next-period employment/income
predictor_matrix_mcs["econ_dist_0", c("econ_emp_bin_1", "log_income_1")] <- 0
predictor_matrix_mcs["econ_dist_1", c("econ_emp_bin_2", "log_income_2")] <- 0

# MCS at t predicts employment at t+1
predictor_matrix_mcs["econ_emp_bin_1", "sf12mcs_dv_0"] <- 1
predictor_matrix_mcs["econ_emp_bin_2", "sf12mcs_dv_1"] <- 1

predictor_matrix_mcs["regime", ] <- 1

# Strata variables does not predict regime
predictor_matrix_mcs["age_group", "regime"] <- 0
predictor_matrix_mcs["edu_group", "regime"] <- 0
```

## Stratified gFormulaMI runs

```{r gformula-mcs}
trt_vars_3 <- c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2")

strata_spec <- list(
  list(col = "sex_dv_base",
       val = "Female",         
       label = "Sex",
       stratum = "Female",
       cache = "em_mcs_sex_female.rds",
       datasets = 200),
  list(col = "sex_dv_base",
       val = "Male",            
       label = "Sex",
       stratum = "Male",
       cache = "em_mcs_sex_male.rds",
       datasets = 200),
  list(col = "age_group",
       val = "Younger (25-40)", 
       label = "Age group",
       stratum = "Younger (25-40)",
       cache = "em_mcs_age_young.rds",
       datasets = 200),
  list(col = "age_group",
       val = "Older (41-65)", 
       label = "Age group",
       stratum = "Older (41-65)",
       cache = "em_mcs_age_older.rds",
       datasets = 200),
  list(col = "edu_group",
       val = "Higher",
       label = "Education",
       stratum = "Higher",
       cache = "em_mcs_edu_higher.rds",
       datasets = 200),
  list(col = "edu_group",
       val = "Lower",
       label = "Education",
       stratum = "Lower",
       cache = "em_mcs_edu_lower.rds",
       datasets = 200)
)

results_mcs <- map(strata_spec, ~ run_em_stratum(
  mids_full       = wide_mids_mcs,
  col             = .x$col,
  val             = .x$val,
  predictor_matrix = predictor_matrix_mcs,
  outcome_var     = "sf12mcs_dv",
  strata_label    = .x$label,
  stratum_label   = .x$stratum,
  trt_vars        = trt_vars_3,
  regimes         = intervention_pattern_3,
  regime_labels   = regimes_3,
  cache_path      = here::here("output", .x$cache),
  datasets        = .x$datasets
))

names(results_mcs) <- map_chr(strata_spec, ~ paste0(.x$label, "_", .x$stratum))
```

## Combine MCS results

```{r combine-mcs}
mcs_marginal <- map_df(results_mcs, "marginal") |>
  mutate(outcome = "Mental Component Score (MCS)")

mcs_diff <- map_df(results_mcs, "diff") |>
  mutate(outcome = "Mental Component Score (MCS)")

saveRDS(mcs_diff,     here::here("output", "em_mcs_diff.rds"))
saveRDS(mcs_marginal, here::here("output", "em_mcs_marginal.rds"))

rm(wide_mids_mcs)
```

# PCS Outcome

```{r wide-pcs}
wide_data_pcs <- long_data |>
  mutate(econ_emp_bin = as_factor(econ_emp_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  race_base,
                  hiqual_dv_base,
                  age_dv_base,
                  age_group,
                  edu_group,
                  sf12pcs_dv_base),
    outcome = sf12pcs_dv,
    mcs_lagged,
    econ_emp_bin,
    log_income,
    lagged_dnc,
    lagged_home_owner,
    lagged_econ_benefits,
    lagged_mastat_dv,
    lagged_age_dv,
    lagged_gor_dv_fact,
    econ_dist,
    static = FALSE,
    waves  = c(0, 1, 2)
  )
```

```{r mice-pcs}
pred_mat_pcs <- make.predictorMatrix(wide_data_pcs)

pred_mat_pcs[, c("lagged_age_dv_0",   "lagged_age_dv_1",   "lagged_age_dv_2")]   <- 0
pred_mat_pcs[, c("lagged_gor_dv_fact_0", "lagged_gor_dv_fact_1", "lagged_gor_dv_fact_2")] <- 0

for (v in c("age_dv_base", "age_group", "edu_group")) {
  pred_mat_pcs[v, ] <- 0
  pred_mat_pcs[, v] <- 0
}

if (file.exists(here::here("output", "wide_mids_pcs_em.rds"))) {
  wide_mids_pcs <- readRDS(here::here("output", "wide_mids_pcs_em.rds"))
  message("Loading cached PCS MIDS (effect modification)...")
} else {
  message("Running MICE for PCS effect modification analysis...")
  wide_mids_pcs <- mice(
    wide_data_pcs,
    predictorMatrix = pred_mat_pcs,
    defaultMethod   = c("norm", "logreg", "pmm", "pmm"),
    m               = 50,
    maxit           = 15,
    seed            = seed
  )
  saveRDS(wide_mids_pcs, here::here("output", "wide_mids_pcs_em.rds"))
}
```

```{r pred-matrix-pcs}
predictor_matrix_pcs <- make_predictor_matrix(wide_data_pcs)

predictor_matrix_pcs[, c("age_dv_base", "age_group", "edu_group")] <- 0

# PCS intermediate values do not predict other intermediates
predictor_matrix_pcs[, "sf12pcs_dv_0"] <- 0
predictor_matrix_pcs[, "sf12pcs_dv_1"] <- 0

# PCS predicted by its own immediate lag, not by baseline PCS
predictor_matrix_pcs["sf12pcs_dv_1", "sf12pcs_dv_0"]    <- 1
predictor_matrix_pcs["sf12pcs_dv_2", "sf12pcs_dv_1"]    <- 1
predictor_matrix_pcs["sf12pcs_dv_1", "sf12pcs_dv_base"] <- 0
predictor_matrix_pcs["sf12pcs_dv_2", "sf12pcs_dv_base"] <- 0

# PCS at t predicts employment and income at t+1
predictor_matrix_pcs["sf12pcs_dv_0", c("econ_emp_bin_1", "log_income_1")] <- 1
predictor_matrix_pcs["sf12pcs_dv_1", c("econ_emp_bin_2", "log_income_2")] <- 1

# Economic distress does not propagate forward
predictor_matrix_pcs["econ_dist_0", c("econ_emp_bin_1", "log_income_1")] <- 0
predictor_matrix_pcs["econ_dist_1", c("econ_emp_bin_2", "log_income_2")] <- 0

# PCS at t predicts employment at t+1
predictor_matrix_pcs["econ_emp_bin_1", "sf12pcs_dv_0"] <- 1
predictor_matrix_pcs["econ_emp_bin_2", "sf12pcs_dv_1"] <- 1

predictor_matrix_pcs["regime", ] <- 1
```

```{r gformula-pcs}
strata_spec_pcs <- map(strata_spec, ~ modifyList(.x, list(
  cache = sub("mcs", "pcs", .x$cache)
)))

results_pcs <- map(strata_spec_pcs, ~ run_em_stratum(
  mids_full        = wide_mids_pcs,
  col              = .x$col,
  val              = .x$val,
  predictor_matrix = predictor_matrix_pcs,
  outcome_var      = "sf12pcs_dv",
  strata_label     = .x$label,
  stratum_label    = .x$stratum,
  trt_vars         = trt_vars_3,
  regimes          = intervention_pattern_3,
  regime_labels    = regimes_3,
  cache_path       = here::here("output", .x$cache),
  datasets         = .x$datasets
))

names(results_pcs) <- map_chr(strata_spec_pcs, ~ paste0(.x$label, "_", .x$stratum))
```

```{r combine-pcs}
pcs_marginal <- map_df(results_pcs, "marginal") |>
  mutate(outcome = "Physical Component Score (PCS)")

pcs_diff <- map_df(results_pcs, "diff") |>
  mutate(outcome = "Physical Component Score (PCS)")

saveRDS(pcs_diff,     here::here("output", "em_pcs_diff.rds"))
saveRDS(pcs_marginal, here::here("output", "em_pcs_marginal.rds"))

rm(wide_mids_pcs)
```

# Results

## Interaction tests

For each effect modifier and outcome, the interaction is tested on the **difference on causal mean differences** contrast per strata (named $\hat{\delta}$).
The test statistic is the difference of the stratum-specific effect estimates divided by the pooled standard error, assuming independence between strata (non-overlapping subgroups).

$$
z = \frac{\hat{\delta}_A - \hat{\delta}_B}{\sqrt{\text{SE}_A^2 + \text{SE}_B^2}}
$$

Where $\hat{\delta}_A$ and $\hat{\delta}_B$ are the causal mean differences for strata A and B, and $\text{SE}_A$ and $\text{SE}_B$ are their respective standard errors. The p-value for the interaction is calculated as $p_{\text{int}} = 2 \times P(|Z| > |z|)$, where $Z$ follows a standard normal distribution under the null hypothesis of no interaction. As we are conducting multiple interaction tests (6 tests: 3 effect modifiers × 2 outcomes), we will apply a Bonferroni correction to adjust the significance threshold to $\alpha = 0.05 / 21 \approx 0.0024$.

```{r interaction-test}
#| output: true

compute_interaction_test <- function(diff_df, strata_var_filter, outcome_label) {

  unique_strata <- diff_df |>
    pull(stratum) |>
    unique()

  diff_df |>
    filter(strata_var == strata_var_filter, intervention != "0-0-0") |>
    arrange(stratum) |>
    reframe(
      outcome   = outcome_label,
      strata_var = strata_var_filter,
      stratum_a  = unique_strata[1],
      effect_a   = mi_effect[1],
      se_a       = mi_se[1],
      stratum_b  = unique_strata[2],
      effect_b   = mi_effect[2],
      se_b       = mi_se[2],
      delta      = effect_a - effect_b,
      se_delta   = sqrt(se_a^2 + se_b^2),
      ci_delta_l = delta - 1.96 * se_delta,
      ci_delta_u = delta + 1.96 * se_delta,
      z          = delta / se_delta,
      p_int      = 2 * pnorm(-abs(z)),
      p_int_adj  = p.adjust(
        p_int,
        method = "bonferroni",
        n = bonferroni_interaction_tests
      ) # total interaction tests across non-reference strategies, effect modifiers, and outcomes
    )
}

all_diff <- bind_rows(mcs_diff, pcs_diff)

treat_strategies <- all_diff |> pull(intervention) |> unique() |> setdiff("0-0-0")
effect_modifier_vars <- c("Sex", "Age group", "Education")
outcome_labels <- c("MCS", "PCS")
bonferroni_interaction_tests <- length(treat_strategies) * length(effect_modifier_vars)

# testing interaction per each treatment strategy with multiple correction for multiple comparisons (6 tests per outcome)
interaction_results <- bind_rows(
  map(
    treat_strategies,
    ~ {
        strategy_diff <- all_diff |> filter(intervention == .x)

        bind_rows(
          compute_interaction_test(strategy_diff, "Sex",       "MCS"),
          compute_interaction_test(strategy_diff, "Age group", "MCS"),
          compute_interaction_test(strategy_diff, "Education", "MCS"),
          compute_interaction_test(strategy_diff, "Sex",       "PCS"),
          compute_interaction_test(strategy_diff, "Age group", "PCS"),
          compute_interaction_test(strategy_diff, "Education", "PCS")
        )
      }
  )
)

interaction_results |>
  select(outcome, strata_var, stratum_a, effect_a, se_a,
         stratum_b, effect_b, se_b, delta, se_delta, ci_delta_l, ci_delta_u, z, p_int, p_int_adj) |>
  arrange(outcome, strata_var) |>
  knitr::kable(digits = 3,
               caption = "Interaction tests across all treatment strategies")
```

## Effect estimates by stratum

```{r tab-diff}
#| output: true

all_diff |>
  filter(intervention != "0-0-0") |>
  mutate(
    n_unemp      = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))),
    intervention = str_replace_all(intervention, c("0" = "E", "1" = "U"))
  ) |>
  select(outcome, strata_var, stratum, intervention, n_unemp,
         mi_effect, mi_se, mi_ll, mi_ul) |>
  arrange(outcome, strata_var, stratum, n_unemp) |>
  knitr::kable(digits = 3,
               caption = "Estimated mean difference vs always-employed (E-E-E), by stratum")
```

## Forest plot: always-unemployed vs always-employed contrast

Probe for effect modification by visualising the causal mean difference for always unemployed (1-1-1) vs always employed (0-0-0) across strata and outcomes. The point estimates and confidence intervals are plotted for MCS and PCS to facilitate comparison of effect modification patterns across outcomes.

```{r graph-em}
#| output: true

graph_em <- all_diff |>
  filter(intervention == "1-1-1") |>
  mutate(
    strata_var = factor(strata_var, levels = c("Sex", "Age group", "Education")),
    stratum    = fct_rev(factor(stratum))
  ) |>
  ggplot(aes(mi_effect, stratum, xmin = mi_ll, xmax = mi_ul, colour = strata_var)) +
  geom_point(size = 2.5) +
  geom_errorbar(width = 0.25) +
  facet_grid(strata_var ~ outcome, scales = "free_y", space = "free_y") +
  scale_colour_manual(values = unname(paletteMartin)[1:3], guide = "none") +
  labs(
    x       = "Estimated mean difference (always unemployed vs always employed)",
    y       = NULL
  ) +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 0))

print(graph_em)

ggsave(graph_em,
       file   = here::here("figs", "graph_em_always_unemployed.png"),
       dpi    = 300,
       width  = 9,
       height = 5)
```

## Full regime plot by stratum

Plots of all 8 intervention strategies (regimes) by stratum and outcome. This allows us to explore whether the effect modification patterns observed in the always-unemployed vs always-employed contrast are consistent across other contrasts (e.g., dose-response patterns, effect of single vs multiple unemployment periods). 

```{r graph-em-full}
#| output: true

graph_em_full <- all_diff |>
  filter(intervention != "0-0-0") |>
  mutate(
    n_unemp    = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |> factor(),
    intervention = str_replace_all(intervention, c("0" = "E", "1" = "U")),
    strata_var = factor(strata_var, levels = c("Sex", "Age group", "Education"))
  ) |>
  ggplot(aes(mi_effect, intervention,
             xmin = mi_ll, xmax = mi_ul,
             colour = stratum, shape = stratum)) +
  geom_point(size = 1.8, position = position_dodge(width = 0.6)) +
  geom_errorbar(width = 0.25, position = position_dodge(width = 0.6)) +
  facet_grid(strata_var ~ outcome, scales = "free_x") +
  scale_colour_manual(
    values = c(
      "Female"          = unname(paletteMartin)[1],
      "Male"            = unname(paletteMartin)[2],
      "Younger (25-40)" = unname(paletteMartin)[3],
      "Older (41-65)"   = unname(paletteMartin)[4],
      "Higher"          = unname(paletteMartin)[5],
      "Lower"           = unname(paletteMartin)[6]
    ),
    name = "Stratum"
  ) +
  scale_shape_manual(
    values = c(
      "Female"          = 16L, "Male"            = 17L,
      "Younger (25-40)" = 16L, "Older (41-65)"   = 17L,
      "Higher"          = 16L, "Lower"           = 17L
    ),
    name = "Stratum"
  ) +
  labs(
    x       = "Estimated mean difference vs always employed (E-E-E)",
    y       = "Intervention strategy"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    strip.text.y    = element_text(angle = 0)
  )

print(graph_em_full)

ggsave(graph_em_full,
       file   = here::here("figs", "graph_em_full_regimes.png"),
       dpi    = 300,
       width  = 10,
       height = 8)
```
