---
title: Financial distress as a mediator of the relationship between employment and health outcomes
author: "Darwin Del Castillo"
format:
  html:
    toc: true
    fig-format: svg
execute:
  output: false
---
# Mediation analysis of employment → mental health via income and financial distress
Randomised Interventional Effect (RIE) mediation via income and financial distress. Decomposes the employment → MCS effect into:
    - IDE = direct effect not operating through income
    - IIE = indirect effect through income (including Inc → Dis → MCS)
    As per the DAGMH in fnct/dags.R:
```{r DAG}
source(here::here("fnct", "dags.R"))
```
## RIE identification (VanderWeele & Tchetgen 2014; Lin et al. 2017):
   (i)   No unmeasured Emp–MCS confounding given C
   (ii*) No unmeasured Inc–MCS confounding given Emp, C, L
   (iii) No unmeasured Emp–Inc confounding given C
Does NOT require the cross-world independence assumption.
Identified with MCS_t → Emp_{t+1} feedback present in this DAG.

```{r setup, include=FALSE}
pacman::p_load(data.table, here, mice, gFormulaMI, magrittr, tidyverse)

# Source functions
source(here::here("fnct", "import_and_clean_data.R"))
source(here::here("fnct", "helpers.R"))

set_seed <- 12345

# ── DATA PIPELINE (identical to 04_with_age.R) ───────────────────────────────
data_clean <- clean_data() |> preproc_data()

eligible_pidp <- intersect(
  data_clean[wave == 2 & response == 1, pidp],
  data_clean[wave == 5 & response == 1, pidp]
)

pop_DT <- data_clean[pidp %in% eligible_pidp]

cols_keep <- c("pidp", "wave", "t0", "sf12mcs_dv", "econ_emp_bin",
               "sf12mcs_dv_base", "log_income", "econ_dist", "age_dv")
long_data <- pop_DT[, ..cols_keep]

long_data <- long_data[, `:=`(t0 = t0 - 2L)][wave %in% 3:5]

no_mh_outcomes <- long_data[,
  .(n_missing = sum(is.na(sf12mcs_dv)), base_missing = all(is.na(sf12mcs_dv_base))),
  by = pidp
][n_missing >= 2 | base_missing == TRUE, .(pidp)]

long_data <- long_data[!no_mh_outcomes, on = .(pidp)]

# matrix of possible intervention patterns for 3 time points (0: employed, 1: unemployed)
intervention_pattern_3 <- asplit(as.matrix(CJ(0:1, 0:1, 0:1)), 1)

wide_data_mcs <- long_data |>
  make_wide(
    id_col    = pidp,
    time_col  = t0,
    base_cols = c(sf12mcs_dv_base),
    outcome   = sf12mcs_dv,
    econ_emp_bin,
    log_income,
    econ_dist,
    age_dv,
    waves     = c(0, 1, 2)
  )

# ── MICE imputation (with caching) ───────────────────────────────────────────
# age_dv is deterministic (increases by 1 each wave) — zero its predictor columns
matrix_imp <- mice::make.predictorMatrix(wide_data_mcs)
matrix_imp[, c("age_dv_0", "age_dv_1", "age_dv_2")] <- 0

mids_path <- here::here("data", "wide_mids_mcs_05.rds")

if (file.exists(mids_path)) {
  wide_mids_mcs <- readRDS(mids_path)
  message("Loaded MICE object from cache: ", mids_path)
} else {
  wide_mids_mcs <- mice::mice(
    wide_data_mcs,
    defaultMethod  = c("norm", "logreg", "pmm", "pmm"),
    m              = 50,
    maxit          = 10,
    seed           = set_seed,
    predictorMatrix = matrix_imp,
    printFlag      = TRUE
  )
  saveRDS(wide_mids_mcs, mids_path)
  message("MICE imputation complete, saved to: ", mids_path)
}

# ── PREDICTOR MATRIX ─────────────────────
predictor_matrix_mcs <- make_predictor_matrix(wide_data_mcs)

## sf12mcs_dv_base does not predict anything beyond time 0
predictor_matrix_mcs[, "sf12mcs_dv_base"] <- 0
predictor_matrix_mcs["regime",         "sf12mcs_dv_base"] <- 1
predictor_matrix_mcs["sf12mcs_dv_0",   "sf12mcs_dv_base"] <- 1
predictor_matrix_mcs["log_income_0",   "sf12mcs_dv_base"] <- 1
predictor_matrix_mcs["econ_dist_0",    "sf12mcs_dv_base"] <- 1
predictor_matrix_mcs["econ_emp_bin_0", "sf12mcs_dv_base"] <- 1

# MCS: Markov chain — only depends on its immediate predecessor
predictor_matrix_mcs["sf12mcs_dv_1", "sf12mcs_dv_0"] <- 1
predictor_matrix_mcs["sf12mcs_dv_2", "sf12mcs_dv_1"] <- 1

# MCS in period t does not predict economic distress in t+1
predictor_matrix_mcs["sf12mcs_dv_0", "econ_dist_1"] <- 0
predictor_matrix_mcs["sf12mcs_dv_1", "econ_dist_2"] <- 0

# econ_dist in period t predicts econ_dist in t+1
predictor_matrix_mcs["econ_dist_1", "econ_dist_0"] <- 1
predictor_matrix_mcs["econ_dist_2", "econ_dist_1"] <- 1

# time-varying covariates at t predict the same covariate at t+1
predictor_matrix_mcs["log_income_1",   "log_income_0"]   <- 1
predictor_matrix_mcs["log_income_2",   "log_income_1"]   <- 1
predictor_matrix_mcs["econ_emp_bin_1", "econ_emp_bin_0"] <- 1
predictor_matrix_mcs["econ_emp_bin_2", "econ_emp_bin_1"] <- 1

# age_dv at t predicts mcs, econ distress, log income, and employment at the same t
predictor_matrix_mcs[, "age_dv_0"] <- 0
predictor_matrix_mcs[, "age_dv_1"] <- 0
predictor_matrix_mcs[, "age_dv_2"] <- 0
predictor_matrix_mcs["regime",         "age_dv_0"] <- 1
predictor_matrix_mcs["regime",         "age_dv_1"] <- 1
predictor_matrix_mcs["regime",         "age_dv_2"] <- 1
predictor_matrix_mcs["sf12mcs_dv_0",   "age_dv_0"] <- 1
predictor_matrix_mcs["sf12mcs_dv_1",   "age_dv_1"] <- 1
predictor_matrix_mcs["sf12mcs_dv_2",   "age_dv_2"] <- 1
predictor_matrix_mcs["econ_dist_0",    "age_dv_0"] <- 1
predictor_matrix_mcs["econ_dist_1",    "age_dv_1"] <- 1
predictor_matrix_mcs["econ_dist_2",    "age_dv_2"] <- 1
predictor_matrix_mcs["log_income_0",   "age_dv_0"] <- 1
predictor_matrix_mcs["log_income_1",   "age_dv_1"] <- 1
predictor_matrix_mcs["log_income_2",   "age_dv_2"] <- 1
predictor_matrix_mcs["econ_emp_bin_0", "age_dv_0"] <- 1
predictor_matrix_mcs["econ_emp_bin_1", "age_dv_1"] <- 1
predictor_matrix_mcs["econ_emp_bin_2", "age_dv_2"] <- 1

# age_dv predicts itself across waves
predictor_matrix_mcs["age_dv_1", "age_dv_0"] <- 1
predictor_matrix_mcs["age_dv_2", "age_dv_1"] <- 1

regimes_mcs <- intervention_pattern_3

regimes_3 <- tibble(intervention = regimes_mcs |>
  map(paste, collapse = "-") |>
  reduce(c))

# ── TOTAL EFFECT ──────────────────────────────────────────────────────────────
# Standard gFormulaImpute with all variables (including income) free to respond
# to the employment regime. Gives E[MCS₂(a)] for each regime.
imps_te <- gFormulaImpute(
  wide_mids_mcs,
  trtVars         = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2"),
  trtRegimes      = regimes_mcs,
  predictorMatrix = predictor_matrix_mcs,
  silent          = TRUE
)

fits_te <- imps_te %$%
  lm(sf12mcs_dv_2 ~ 0 + factor(regime))

te_marginal <- syntheticPool(fits_te) |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(
    te_effect = Estimate,
    te_se     = sqrt(Total),
    te_ll     = `95% CI L`,
    te_ul     = `95% CI U`
  ) |>
  bind_cols(regimes_3)

# ── RIE VIA INCOME: STAGE 1 ─────────────────────────────────────────────────
# Reference regime (all employed = 0,0,0), income is free to re-impute.
# The resulting log_income values ARE G_{a*|C}: draws from the income distribution
# under reference employment, conditional on baseline confounders C.
ref_imps <- gFormulaImpute(
  wide_mids_mcs,
  trtVars         = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2"),
  trtRegimes      = c(0, 0, 0),         # plain vector: reference regime
  predictorMatrix = predictor_matrix_mcs,
  silent          = TRUE
)

# ── RIE VIA INCOME: STAGE 2 ──────────────────────────────────────────────────
# All 8 employment regimes; income will be frozen post-hoc to Stage 1 draws,
# breaking the Emp → Inc → MCS pathway while keeping the Emp → MCS direct path.
rie_imps <- gFormulaImpute(
  wide_mids_mcs,
  trtVars         = c("econ_emp_bin_0", "econ_emp_bin_1", "econ_emp_bin_2"),
  trtRegimes      = regimes_mcs,
  predictorMatrix = predictor_matrix_mcs,
  silent          = TRUE
)

# Sanity check: verify imputation counts
M             <- wide_mids_mcs$m
n_ref_per_imp <- nrow(mice::complete(ref_imps, action = 1))
n_rie_per_imp <- nrow(mice::complete(rie_imps, action = 1))
n_regimes     <- n_rie_per_imp / n_ref_per_imp
stopifnot(n_regimes == length(regimes_mcs))  # must equal 8

# ── Outcome model formula (read from predictor matrix) ────────────────────────
# gFormulaImpute already generated sf12mcs_dv_2 using natural mediator values.
# To apply the RIE correctly we must RE-PREDICT sf12mcs_dv_2 after freezing the
# mediator, using the outcome model fitted on each MI-imputed dataset.
# Predictors of sf12mcs_dv_2 per predictor_matrix_mcs (sliding window + Markov override):
#   sf12mcs_dv_1, econ_emp_bin_2, log_income_2, econ_dist_2, age_dv_2
outcome_preds   <- setdiff(names(which(predictor_matrix_mcs["sf12mcs_dv_2", ] == 1)), "regime")
outcome_formula <- reformulate(outcome_preds, response = "sf12mcs_dv_2")

# For each imputation k:
#   1. Fit outcome model on the MI-complete dataset (captures natural associations)
#   2. Freeze income in the synthetic data (tiled across 8 regime blocks)
#   3. Re-predict sf12mcs_dv_2 using frozen income + Stage 2 other predictors
#   4. Fit marginal mean model on the counterfactual outcome
fits_rie_analyses <- lapply(seq_len(M), function(k) {
  ref_k  <- mice::complete(ref_imps,    action = k)
  rie_k  <- mice::complete(rie_imps,    action = k)
  imp_k  <- mice::complete(wide_mids_mcs, action = k)
  out_mod <- lm(outcome_formula, data = imp_k)
  rie_k$log_income_0 <- rep(ref_k$log_income_0, times = n_regimes)
  rie_k$log_income_1 <- rep(ref_k$log_income_1, times = n_regimes)
  rie_k$log_income_2 <- rep(ref_k$log_income_2, times = n_regimes)
  rie_k$sf12mcs_dv_2 <- predict(out_mod, newdata = rie_k)   # counterfactual outcome
  lm(sf12mcs_dv_2 ~ 0 + factor(regime), data = rie_k)
})

fits_rie <- list(analyses = fits_rie_analyses)
mi_rie   <- syntheticPool(fits_rie)

rie_pooled <- mi_rie |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(
    rie_effect = Estimate,
    rie_se     = sqrt(Total),
    rie_ll     = `95% CI L`,
    rie_ul     = `95% CI U`
  ) |>
  bind_cols(regimes_3)

# ── DECOMPOSITION: TE, IDE, IIE ──────────────────────────────────────────────
# All effects contrasted against the reference regime: always employed (0-0-0).
#
#   TE  = E[MCS₂(a)] − E[MCS₂(a*)]          : total effect of employment on MCS
#   IDE = Q(a, a*) − Q(a*, a*)               : direct effect (not through income)
#   IIE = TE − IDE                           : indirect effect via income
#                                              (includes Inc→MCS and Inc→Dis→MCS)
ref_label <- "0-0-0"
te_ref    <- te_marginal$te_effect[te_marginal$intervention == ref_label]
rie_ref   <- rie_pooled$rie_effect[rie_pooled$intervention == ref_label]

rie_decomposition <- rie_pooled |>
  left_join(
    te_marginal |> select(intervention, te_effect),
    by = "intervention"
  ) |>
  mutate(
    TE  = te_effect - te_ref,
    IDE = rie_effect - rie_ref,
    IIE = TE - IDE
  ) |>
  select(intervention, TE, IDE, IIE, rie_effect, rie_se, rie_ll, rie_ul)

# ── TABLE ─────────────────────────────────────────────────────────────────────
# Display-ready version: exclude reference regime, format strategy labels,
# add proportion mediated column.
tab_display <- rie_decomposition |>
  filter(intervention != ref_label) |>
  mutate(
    Intervention   = str_replace_all(intervention, c("0" = "E", "1" = "U")),
    `% via income` = round(IIE / TE * 100, 1)
  ) |>
  select(Intervention, TE, IDE, IIE, `% via income`)

# Print to console
tab_display |>
  knitr::kable(
    digits  = 3,
    caption = "RIE decomposition: Employment -> MCS via income (E=employed, U=unemployed)"
  )

# Save to figs/ as PNG for inclusion in slides
tab_display |>
  knitr::kable(
    digits     = 3,
    col.names  = c("Intervention", "TE", "IDE (not via income)", "IIE (via income)", "% mediated"),
    caption    = "Decomposition of employment effect on MCS-12 via income (RIE)"
  ) |>
  kableExtra::kable_styling(
    full_width        = FALSE,
    font_size         = 14,
    bootstrap_options = c("striped", "hover")
  ) |>
  kableExtra::add_footnote(
    "E = employed, U = unemployed. Reference: E-E-E (always employed).
     TE = total effect; IDE = interventional direct effect;
     IIE = interventional indirect effect via income (Inc→MCS and Inc→Dis→MCS).",
    notation = "none"
  ) |>
  kableExtra::save_kable(
    file = here::here("figs", "table_rie_decomposition.png"),
    zoom = 2
  )

# ── GRAPH ─────────────────────────────────────────────────────────────────────
graph_rie <- rie_decomposition |>
  filter(intervention != ref_label) |>
  select(intervention, TE, IDE, IIE) |>
  pivot_longer(c(TE, IDE, IIE), names_to = "Effect", values_to = "Estimate") |>
  mutate(
    n_unemp      = map_int(intervention, ~ length(str_extract_all(.x, "1", simplify = TRUE))) |>
      factor(),
    intervention = str_replace_all(intervention, c("0" = "E", "1" = "U")),
    Effect       = factor(Effect, levels = c("TE", "IDE", "IIE"))
  ) |>
  ggplot(aes(Estimate, intervention, colour = Effect, shape = Effect)) +
  geom_point(size = 2.5, position = position_dodge(width = 0.6)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.4) +
  facet_wrap(~ n_unemp, ncol = 1,
             labeller = labeller(n_unemp = ~ paste("Unemployed periods:", .x))) +
  scale_colour_brewer(palette = "Set1",
                      labels  = c("Total (TE)", "Direct — not via income (IDE)",
                                  "Indirect — via income (IIE)")) +
  scale_shape_manual(values = c(16, 17, 15),
                     labels = c("Total (TE)", "Direct — not via income (IDE)",
                                "Indirect — via income (IIE)")) +
  labs(
    x      = "Effect on MCS-12 mental health score",
    y      = "Intervention strategy (E = employed, U = unemployed)",
    title  = "Decomposition of employment effect on MCS via income",
    colour = "Effect type",
    shape  = "Effect type"
  )

print(graph_rie)

# ══════════════════════════════════════════════════════════════════════════════
# RIE VIA FINANCIAL DISTRESS
# ══════════════════════════════════════════════════════════════════════════════
# Reuses ref_imps and rie_imps from the income section (same gFormulaImpute
# runs). Only the frozen variable changes: econ_dist instead of log_income.
#
# DAG pathway captured by IIE_fd:
#   Emp_t → Dis_t → MCS_t           (direct distress path)
#   Emp_t → Inc_t → Dis_t → MCS_t   (income-to-distress path)
#   Emp_t → Dis_t → Dis_{t+1} → MCS_{t+1}  (distress autocorrelation)
#
# IDE_fd captures everything NOT operating through financial distress:
#   direct Emp → MCS, and Emp → Inc → MCS (income not via distress).

fits_fd_analyses <- lapply(seq_len(M), function(k) {
  ref_k   <- mice::complete(ref_imps,     action = k)
  rie_k   <- mice::complete(rie_imps,     action = k)
  imp_k   <- mice::complete(wide_mids_mcs, action = k)
  out_mod <- lm(outcome_formula, data = imp_k)
  # Freeze financial distress at all 3 time points to reference-regime draws
  rie_k$econ_dist_0 <- rep(ref_k$econ_dist_0, times = n_regimes)
  rie_k$econ_dist_1 <- rep(ref_k$econ_dist_1, times = n_regimes)
  rie_k$econ_dist_2 <- rep(ref_k$econ_dist_2, times = n_regimes)
  rie_k$sf12mcs_dv_2 <- predict(out_mod, newdata = rie_k)   # counterfactual outcome
  lm(sf12mcs_dv_2 ~ 0 + factor(regime), data = rie_k)
})

fits_fd <- list(analyses = fits_fd_analyses)
mi_fd   <- syntheticPool(fits_fd)

fd_pooled <- mi_fd |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(
    fd_effect = Estimate,
    fd_se     = sqrt(Total),
    fd_ll     = `95% CI L`,
    fd_ul     = `95% CI U`
  ) |>
  bind_cols(regimes_3)

# ── DECOMPOSITION: TE, IDE_fd, IIE_fd ────────────────────────────────────────
fd_ref <- fd_pooled$fd_effect[fd_pooled$intervention == ref_label]

fd_decomposition <- fd_pooled |>
  left_join(
    te_marginal |> select(intervention, te_effect),
    by = "intervention"
  ) |>
  mutate(
    TE     = te_effect - te_ref,
    IDE_fd = fd_effect - fd_ref,
    IIE_fd = TE - IDE_fd
  ) |>
  select(intervention, TE, IDE_fd, IIE_fd, fd_effect, fd_se, fd_ll, fd_ul)

# ── TABLE ─────────────────────────────────────────────────────────────────────
tab_fd <- fd_decomposition |>
  filter(intervention != ref_label) |>
  mutate(
    Intervention = str_replace_all(intervention, c("0" = "E", "1" = "U")),
    `% via FD`   = round(IIE_fd / TE * 100, 1)
  ) |>
  select(Intervention, TE, IDE_fd, IIE_fd, `% via FD`)

# Print to console
tab_fd |>
  knitr::kable(
    digits  = 3,
    caption = "RIE decomposition: Employment -> MCS via financial distress (E=employed, U=unemployed)"
  )

# Save to figs/ as PNG for inclusion in slides
tab_fd |>
  knitr::kable(
    digits    = 3,
    col.names = c("Intervention", "TE", "IDE (not via FD)", "IIE (via FD)", "% mediated"),
    caption   = "Decomposition of employment effect on MCS-12 via financial distress (RIE)"
  ) |>
  kableExtra::kable_styling(
    full_width        = FALSE,
    font_size         = 14,
    bootstrap_options = c("striped", "hover")
  ) |>
  kableExtra::add_footnote(
    "E = employed, U = unemployed. Reference: E-E-E (always employed).
     TE = total effect; IDE = interventional direct effect;
     IIE = interventional indirect effect via financial distress (Emp→Dis→MCS and Emp→Inc→Dis→MCS).",
    notation = "none"
  ) |>
  kableExtra::save_kable(
    file = here::here("figs", "table_rie_decomposition_fd.png"),
    zoom = 2
  )
```